---
title: "Pipeline step 6 genetic principal components"
---


9/29/22 Added the genetic PCs to the pipeline for inclusion as regression covariates. The first 40 are calculated and provided by the UK Biobank in data field 22009 "Genetic principal components". Based on an email conversation with Alex Hatoum we decided to include 10 of the PCs since this is generally enough in a European population.

```{r}
library(dplyr) # Run install.packages("tidyverse") to get this
library(tidyr) # Run install.packages("tidyr") to get this
library(ggplot2)
library("reshape") # Run install.packages("reshape") to get this
library(kableExtra) # Run install.packages("kableExtra")

```


Set the filepath to get input files
```{r}
# Home computer
input_path <- "file_path"

```

Set the filepath to save outputs to
```{r}
# Home computer
output_path <- "file_path"

```

Set the date the pipeline is being run
```{r}

run_date <- "yyyy-mm-dd"

```


Load the PCs
```{r}

genetic_PC <- read.csv(paste0(input_path, "ukb_pc20.csv"))

```

```{r}

load(paste0(output_path, "filtered_eid_p2_", run_date, ".RData"))

genetic_PC <- genetic_PC %>%
  filter(genetic_PC$n_eid %in% filtered_eid)

```


Select the first 10
```{r}

genetic_PC <- genetic_PC[,1:11]

```

Rename the columns so they are more readable
```{r}

genetic_PC <- genetic_PC %>%
  rename_with(~ gsub("n_22009_0_", "", .x)) %>%
  rename_with(~ paste0("PC_", .x)) %>%
  # Fix n_eid, it is not a PC
  dplyr::rename("n_eid" = PC_n_eid)

head(genetic_PC)

```

Remove any rows that are missing genetic components

```{r}

print(paste0("Before: ", nrow(genetic_PC)))
n_before <- nrow(genetic_PC)

genetic_PC <- na.omit(genetic_PC)

print(paste0("After: ", nrow(genetic_PC)))
print(paste0((n_before - nrow(genetic_PC)), " participants dropped for missing genetic PCs"))

```

Save the genetic PCs and a list of the IDs which have PCs
```{r}

genetic_PC_eid <- genetic_PC$n_eid

save(genetic_PC_eid, file = paste0(output_path, "genetic_PC_eid_", Sys.Date(), ".RData"))

save(genetic_PC, file = paste0(output_path, "genetic_PC_", Sys.Date(), ".RData"))

```

