---
title: "Pipeline Step 0: IDP names with identifier columns"
---

We have a file with IDPs and their human readable names. This script cleans that up and makes a tidy data table with info about each IDP which can be joined to future regression results and used for selecting IDPs.

Columns created:
IDP_name: Human readable name of what the IDP represents
IDP: UK Biobank data field for the IDP
numeric_IDP: just the numeric part of the IDP data field
MRI: MRI sequence the IDP is from
map: which parcellation scheme the IDP is part of
dMRI: for dMRI IDPs only, which NODDI measure the IDP represents
dimension: area, thickness, or volume (relevant for T1 IDPs)
region: what region the IDP is in
cortical: is the IDP cortical or subcortical
hemisphere: right or left hemisphere (or neither)
paired: if the IDP is part of a left / right pair, this column contains the corresponding IDP on the opposite hemisphere
display_group: a tag useful for making manhattan plots of the IDPs


```{r}
library(dplyr) # Run install.packages("tidyverse") to get this
library(tidyr) # Run install.packages("tidyr") to get this
library(ggplot2)
library("reshape") # Run install.packages("reshape") to get this
library("broom")
library(kableExtra) # Run install.packages("kableExtra")

```


Set the filepath to get input files
```{r}
# Home computer
input_path <- "file_path"

```

Set the filepath to save outputs to
```{r}
# Home computer
output_path <- "file_path"

```

Import the text file with the IDP names
Why it has this specific format I am not sure.
It is important to use read.delim instead of read.table because otherwise the hashtags break everything

```{r}
# Load IDP names
IDP_names <- read.delim(file = paste0("IDPs_37741_names.txt"), sep = "#", header = FALSE)

```

This imports with 2 columns
V1 contains the IDP identifier which will match to the imaging data
V2 is the human legible name for the region

There is a bunch of stuff along with the IDP name in the first column - strip this out

Implement the string manipulations
Had to replace [][[]] with map and pluck from the purrr package because Tidyverse
https://stackoverflow.com/questions/55663831/how-to-index-the-first-element-of-a-list-and-apply-it-to-each-row-of-a-datafram

```{r}
IDP_names <- IDP_names %>%
  mutate(IDP = strsplit(IDP_names$V1, " ")) %>%
  mutate(IDP = purrr::map(IDP, purrr::pluck, 2)) %>%
  mutate(IDP = gsub(" ", "", paste("X", gsub("-", ".", IDP))))
  
```

Now start creating additional columns based on info in column V2 which is human legible names

Make columns to indicate which MRI modality, and then which segmentation map the IDPs are from
UK Biobank IDP data fields are numbered, make a numeric version of the IDP names to use in sorting
```{r}
IDP_names <- IDP_names %>%
  mutate(numeric_IDP = strsplit(IDP_names$V1, " ")) %>%
  mutate(numeric_IDP = purrr::map(numeric_IDP, purrr::pluck, 2)) %>%
  mutate(numeric_IDP = strsplit(as.character(numeric_IDP), "-")) %>%
  mutate(numeric_IDP = purrr::map(numeric_IDP, purrr::pluck, 1)) %>%
  mutate(numeric_IDP = as.numeric(numeric_IDP))

```

Remove the V1 column since it isn't needed anymore, and also rename V2

```{r}

IDP_names <- IDP_names %>%
  select(! V1) %>%
  dplyr::rename("IDP_name" = V2)

```

Initialize the MRI and map columns

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = NA) %>%
  mutate(map = NA)

```

Regional grey matter (FAST) are fields 25782 to 25920
There are 139 IDP on the showcase, as of 9/8/21 we have all of them in this naming file

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 25782 & numeric_IDP <= 25920, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 25782 & numeric_IDP <= 25920, "FAST", map))

# Check that the correct number are found in the naming file
length(grep("FAST", IDP_names$map))

```

Subcortical volumes (FIRST) are fields 25011 to 25024
There are 14 IDP on the showcase, as of 9/8/21 all are present in this file

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 25011 & numeric_IDP <= 25024, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 25011 & numeric_IDP <= 25024, "FIRST", map))

# Check that the correct number are found in the naming file
length(grep("FIRST", IDP_names$map))

```

Freesurfer ASEG are fields 26501 to 26599

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 26501 & numeric_IDP <= 26599, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 26501 & numeric_IDP <= 26599, "freesurfer_ASEG", map))

# Check that the correct number are found in the naming file
length(grep("freesurfer_ASEG", IDP_names$map))

```

Freesurfer BA exvivo are fields 27059 to 27143
There are 84 IDP listed on the showcase

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 27059 & numeric_IDP <= 27143, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 27059 & numeric_IDP <= 27143, "freesurfer_BA", map))

# Check that the correct number are found in the naming file
length(grep("freesurfer_BA", IDP_names$map))

```

Freesurfer a2009s are fields 27329 to 27772
There are 444 IDP listed on the showcase

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 27329 & numeric_IDP <= 27772, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 27329 & numeric_IDP <= 27772, "freesurfer_a2009s", map))

# Check that the correct number are found in the naming file
length(grep("freesurfer_a2009s", IDP_names$map))

```

Freesurfer DKT IDPs are 27143 to 27328 (the left are assigned before the right, so they alternate counting up in showcase)
The showcase reports 186 IDPs in this set

```{r}

IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 27143 & numeric_IDP <= 27328, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 27143 & numeric_IDP <= 27328, "freesurfer_DKT", map))

# Check that the correct number are found in the naming file
length(grep("freesurfer_DKT", IDP_names$map))

```

Freesurfer desikan gw IDPs are 26989 to 27058
There should be 70

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 26989 & numeric_IDP <= 27058, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 26989 & numeric_IDP <= 27058, "desikan_gw", map))

# Check that the correct number are found in the naming file
length(grep("desikan_gw", IDP_names$map))

```

Freesurfer desikan pial IDPs are 26923 to 26988
There should be 66

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 26923 & numeric_IDP <= 26988, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 26923 & numeric_IDP <= 26988, "desikan_pial", map))

# Check that the correct number are found in the naming file
length(grep("desikan_pial", IDP_names$map))
```

Freesurfer desikan white IDPs are 26721 to 26921
There should be 202

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 26721 & numeric_IDP <= 26922, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 26721 & numeric_IDP <= 26922, "desikan_white", map))

# Check that the correct number are found in the naming file
length(grep("desikan_white", IDP_names$map))
```

Freesurfer subsegmentation IDPs are 26600 to 26720
There should be 121

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 26600 & numeric_IDP <= 26720, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 26600 & numeric_IDP <= 26720, "freesurfer_subseg", map))

# Check that the correct number are found in the naming file
length(grep("freesurfer_subseg", IDP_names$map))

```

Label fields for T2 MRI
Yoonhoo's slide says there should be 33 of these with different subregions
There are two measures under category 112 T2 weighted brain MRI

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = case_when(
    numeric_IDP == 25781 ~ "T2",
    numeric_IDP == 25736 ~ "T2",
    TRUE ~ MRI
  ))

```


There are 18 fields under category 109 susceptibility weighted brain MRI
Some of these fields are 202 series and are not in our imaging names file
Label the ones that are, range is 25026 to 25039, although not all values in between these are present
There are also a few completely outside this range
```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 25026 & numeric_IDP <= 25039, "T2", MRI))

IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP == 25738, "T2", MRI))


# Check that the correct number are found in the naming file
length(grep("T2", IDP_names$MRI))


```
As of 9/8/21 we find 17 total T2 fields through the data set.


Task functional MRI
Yoonhoo's slide lists 12 IDPs for this, the showcase has 28 fields
A few of the fields in the showcase are 202- series, I think this is processed information that would be in with the phenotypes

Two ranges:
25740 to 25768, but not all of these numbers are present
25040 to 25054, not all numbers in between present

9/8/21 we are capturing 25 fields in the naming document, for now this is good enough
```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 25740 & numeric_IDP <= 25768, "task_functional", MRI)) %>%
  mutate(MRI = ifelse(numeric_IDP >= 25040 & numeric_IDP <= 25054, "task_functional", MRI))

# Check that the correct number are found in the naming file
length(grep("task_functional", IDP_names$MRI))
```

Diffusion weighted MRI

dMRI skeleton is IDP fields 25056 to 25487 for 432 data fields
```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 25056 & numeric_IDP <= 25487, "dMRI", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 25056 & numeric_IDP <= 25487, "skeleton", map))

# Check that the correct number are found in the naming file
length(grep("skeleton", IDP_names$map))

```

dMRI weighted means is IDP fields 25488 to 25730 for 243 data fields

```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 25488 & numeric_IDP <= 25730, "dMRI", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 25488 & numeric_IDP <= 25730, "weighted_mean", map))

# Check that the correct number are found in the naming file
length(grep("weighted_mean", IDP_names$map))

```

Within diffusion MRI there are different types, make a column to capture these
```{r}

IDP_names <- IDP_names %>%
  mutate(dMRI = case_when(
    grepl("ean FA", IDP_names$IDP_name) ~ "FA",
    grepl("ean MD", IDP_names$IDP_name) ~ "MD",
    grepl("ean MO", IDP_names$IDP_name) ~ "MO",
    grepl("ean ISOVF", IDP_names$IDP_name) ~ "ISOVF",
    grepl("ean ICVF", IDP_names$IDP_name) ~ "ICVF",
    grepl("ean L1", IDP_names$IDP_name) ~ "L1",
    grepl("ean L2", IDP_names$IDP_name) ~ "L2",
    grepl("ean L3", IDP_names$IDP_name) ~ "L3",
    grepl("ean OD", IDP_names$IDP_name) ~ "OD", # OD is present in the naming file but did not appear on Yoonhoo's slide
  ))


```

Resting functional MRI
Range is between 25739 and 25755, not including all these numbers
I think these are interspersed with the task functional because they were getting classed that way
14 data fields are listed in the showcase, 10 of them should be in the naming set
The other 4 are outside the range of what is included in our IDPs
```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP %in% c(25739, 25741, 25743, 25744, 25750, 25751, 25752, 25753, 25754, 25755),
                      "resting_functional", MRI))

# Check that the correct number are found in the naming file
length(grep("resting_functional", IDP_names$MRI))
```

For the T1 segmentations, add a dimension column to indicate what is being measured (area, thickness, volume)
```{r}

IDP_names <- IDP_names %>%
  mutate(dimension = case_when(
    grepl("area", tolower(IDP_names$IDP_name), fixed = TRUE) ~ "area",
    grepl("thickness", tolower(IDP_names$IDP_name), fixed = TRUE) ~ "thickness",
    grepl("volume", tolower(IDP_names$IDP_name), fixed = TRUE) ~ "volume",
    grepl("mean intensity", tolower(IDP_names$IDP_name), fixed = TRUE) ~ "mean_intensity"))

```

Most IDPs are assigned to a group, check what is remaining

```{r}

no_MRI <- IDP_names %>%
  filter(is.na(MRI))

no_MRI

```

Global (total) brain volumes

25000 to 25010 are total brain volumes derived from T1. For map I put global
As is 25025 (volume of brain stem + 4th ventricle)
25731 to 25738 are for quality control and processing, call the MRI on these QC

The remaining IDP is acquisition protocol phase, also for QC and processing
```{r}
IDP_names <- IDP_names %>%
  mutate(MRI = ifelse(numeric_IDP >= 25000 & numeric_IDP <= 25010, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP >= 25000 & numeric_IDP <= 25010, "global", map)) %>%
  mutate(MRI = ifelse(numeric_IDP == 25025, "T1", MRI)) %>%
  mutate(map = ifelse(numeric_IDP == 25025, "global", map)) %>%
  mutate(MRI = ifelse(numeric_IDP >= 25731 & numeric_IDP <= 25738, "QC", MRI)) %>%
  mutate(MRI = ifelse(numeric_IDP == 25780, "QC", MRI))

```


Add a column to express when structures are part of a superstructure - ex. amygdala is part of the limbic system
10/31/21 Updated this so lobes are not a separate column

Import the table with names assigned by Yoonhoo that was emailed out on 11/30/21
This was part of a larger file with her results, I pulled out the names to a separate file to use in this analysis


```{r}

region_mapping <- read.csv(paste0("Yoonhoo_assigned_IDP_regions_11-30-21.csv"))

```

Select out the IDP and corresponding region name, then left join into the IDP_names
I was previously using word search to assign these but the by hand method is necessary to ensure we made the right decision about each one.
```{r}

region_mapping <- region_mapping %>%
  select(IDP, Region) %>%
  dplyr::rename("numeric_IDP" = IDP, "region" = Region)
  

IDP_names <- left_join(IDP_names, region_mapping, by = "numeric_IDP")

```

There are some measures with no region, some of these are global measures but not all
Label measures which reflect global volumes with region = global
```{r}

IDP_names <- IDP_names %>%
  mutate(region = ifelse(numeric_IDP >= 25000 & numeric_IDP <= 25010, "global", region)) %>%
  mutate(region = ifelse(numeric_IDP == 25025, "global", region))

```

Measures of global volume from Freesurfer ASEG
Ensure these have global as the region for these so they can be easily identified
26514 = ASEG total volume
26518 = ASEG total grey matter volume
26527 = ASEG CSF volume
26553, 26584 = left and right white matter volume

IDPs 26755 and 26856 represent left and right "mean thickness of global mean thickness". We use them as a global variable alongside the aseg global variables
```{r}

IDP_names <- IDP_names %>%
  mutate(region = ifelse(numeric_IDP %in% c(25001, 25003, 25005, 25007, 25009,
                                            26514, 26518, 26527, 26528, 26553, 26584), "global", region))

```

For controlling in models and some results we want total white matter volume and an overall mean thickness
We also calculate overall FA, MD, ICVF, ISOVF
Make rows to accommodate these IDP derived measures
```{r}

derived_global <- data.frame(matrix(ncol = ncol(IDP_names), nrow = 14))
colnames(derived_global) <- colnames(IDP_names)

derived_global$IDP_name <- c("total white matter", "total cortical volume", "global mean thickness",
                             "fast_cortical", "fast_subcortical", "scaled_fast_cortical", "scaled_fast_subcortical",
                             "total_cerebellum",
                             "mean FA", "mean MD", "mean ICVF", "mean ISOVF", "mean MO", "mean OD")
derived_global$IDP <- c("total_wm", "total_cortex", "global_mean_thk",
                        "fast_cortical", "fast_subcortical", "scaled_fast_cortical", "scaled_fast_subcortical",
                        "total_cerebellum",
                        "mean_FA", "mean_MD", "mean_ICVF", "mean_ISOVF", "mean_MO", "mean_OD")
derived_global$MRI <- c(rep("T1", 8), rep("dMRI", 6))
derived_global$dMRI <- c(rep(NA, 8), "FA", "MD", "ICVF", "ISOVF", "MO", "OD")
derived_global$region <- rep("global", 14)
derived_global$map <- c(rep("derived", 14))
derived_global$dimension <- c("volume", "volume", "thickness", rep("volume", 5), rep(NA, 6))


```

Stick the derived global IDPs onto the main table
```{r}

IDP_names <- rbind(IDP_names, derived_global)

```

Freesurfer ASEG has some global / overall measures mixed in with regional, make sure these are all marked
I identified these by reading through the UK Biobank showcase (https://biobank.ndph.ox.ac.uk/showcase/label.cgi?id=190)
```{r}

aseg_global <- c(
  "X26514.2.0", #Volume of BrainSeg (whole brain)
  "X26515.2.0", #	Volume of BrainSegNotVent (whole brain)
  "X26516.2.0", #	Volume of BrainSegNotVentSurf (whole brain)
  "X26527.2.0", #	Volume of CSF (whole brain)
  "X26553.2.0", #	Volume of CerebralWhiteMatter (left hemisphere)
  "X26584.2.0", #	Volume of CerebralWhiteMatter (right hemisphere)
  "X26552.2.0", #	Volume of Cortex (left hemisphere)
  "X26583.2.0", #	Volume of Cortex (right hemisphere)
  "X26521.2.0", #	Volume of EstimatedTotalIntraCranial (whole brain)
  "X26517.2.0", #	Volume of SubCortGray (whole brain)
  "X26519.2.0", #	Volume of SupraTentorial (whole brain)
  "X26520.2.0", #	Volume of SupraTentorialNotVent (whole brain)
  "X26522.2.0", #	Volume of VentricleChoroid (whole brain)
  "X26528.2.0", #	Volume of WM-hypointensities (whole brain)
  "X26529.2.0", #	Volume of non-WM-hypointensities (whole brain)
  "X26566.2.0", #	Volume of vessel (left hemisphere)
  "X26597.2.0") #	Volume of vessel (right hemisphere)

aseg_QC <- c(
  "X26568.2.0", #	Number of HolesBeforeFixing (left hemisphere)
  "X26599.2.0", #	Number of HolesBeforeFixing (right hemisphere)
  "X26536.2.0", #	Volume-ratio of BrainSegVol-to-eTIV (whole brain)
  "X26537.2.0") #	Volume-ratio of MaskVol-to-eTIV (whole brain)

aseg_cerebellum <- c(
  "X26557.2.0", #	Volume of Cerebellum-Cortex (left hemisphere)
  "X26588.2.0", #	Volume of Cerebellum-Cortex (right hemisphere)
  "X26556.2.0", #	Volume of Cerebellum-White-Matter (left hemisphere)
  "X26587.2.0") #	Volume of Cerebellum-White-Matter (right hemisphere)

aseg_ventricles <-c(
  "X26523.2.0", #	Volume of 3rd-Ventricle (whole brain)
  "X26524.2.0", #	Volume of 4th-Ventricle (whole brain)
  "X26525.2.0", #	Volume of 5th-Ventricle (whole brain)
  "X26555.2.0", #	Volume of Inf-Lat-Vent (left hemisphere)
  "X26586.2.0", #	Volume of Inf-Lat-Vent (right hemisphere)
  "X26554.2.0", #	Volume of Lateral-Ventricle (left hemisphere)
  "X26585.2.0") #	Volume of Lateral-Ventricle (right hemisphere)


```

```{r}

IDP_names <- IDP_names %>%
  mutate(region = ifelse(IDP %in% aseg_global, "global", region)) %>%
  mutate(region = ifelse(IDP %in% aseg_QC, "QC", region)) %>%
  mutate(region = ifelse(IDP %in% aseg_cerebellum, "cerebellum", region)) %>%
  mutate(region = ifelse(IDP %in% aseg_ventricles, "ventricles", region)) %>%
  
  # A few ASEG regions also get mis-classified as area, fix this
  mutate(dimension = ifelse(map == "freesurfer_ASEG", 
                            ifelse(dimension != "mean_intensity", "volume", dimension), dimension))

```


Group the regions by cortical or subcortical
```{r}

unique(IDP_names$region)

```

Make a table with all the regions, and where they fall
This is not the most efficient way to make a table, but it is easy to read and edit how the regions are assigned
```{r}

regions_to_cortical <- data.frame((rbind(
  c("caudate", "subcortical"),
  c("global", NA),
  c("thalamus", "subcortical"),
  c("putamen", "subcortical"),
  c("pallidum", "subcortical"),
  c("hippocampus", "subcortical"),
  c("amygdala", "subcortical"),
  c("accumbens", "subcortical"),
  c("peduncle", "subcortical"),
  c("pontine crossing tract", "subcortical"),
  c("corpus callosum", "subcortical"),
  c("fornix", "subcortical"),
  c("corticospinal tract", "subcortical"),
  c("medial lemniscus", "subcortical"),
  c("internal capsule", "subcortical"),
  c("corona radiata", "subcortical"),
  c("sagittal stratum", "subcortical"),
  c("external capsule", "subcortical"),
  c("cingulate   ", "cortical"),
  c("longitudinal fasciculus", "subcortical"),
  c("occipital", "cortical"),
  c("uncinate fasciculus", "subcortical"),
  c("acoustic radiation", "subcortical"),
  c("cingulum", "subcortical"),
  c("forceps major", "subcortical"),
  c("forceps minor", "subcortical"),
  c("frontal", "cortical"),
  c("insula", "cortical"),
  c("precentral", "cortical"),
  c("temporal", "cortical"),
  c("parietal", "cortical"),
  c("SMA", "cortical"),
  c("parahippocampal", "cortical"), # Fixed this 8/16/22
  c("striatum", "subcortical"),
  c("brainstem", "subcortical"),
  c("cerebellum", "subcortical"),
  c("fusiform", "cortical"),
  c("paracentral", "cortical"))))


regions_to_cortical <- regions_to_cortical %>%
  dplyr::rename("region" = X1, "cortical" = X2)

```

Add this as a column to the IDP names
```{r}

IDP_names <- IDP_names %>%
  left_join(regions_to_cortical, by = "region")

```



Make columns to indicate right and left hemispheres
```{r}

IDP_names <- IDP_names %>%
  mutate(hemisphere = ifelse(grepl("right", tolower(IDP_names$IDP_name), fixed = TRUE), "right",
                             ifelse(grepl("left", tolower(IDP_names$IDP_name), fixed = TRUE), "left",
                                    ifelse(grepl("vermis", tolower(IDP_names$IDP_name), fixed = TRUE), "vermis",
                                    NA))))

```


Use left / right to detect paired measures and match them to their opposite
Split on opening parentheses to get rid of the (left) / (right) in the names
Have to add the T1 map to strip name or else there will be indistinguishable duplicates in the name
```{r}

left_hemisphere <- IDP_names %>%
  mutate(strip_name = strsplit(IDP_names$IDP_name, "\\(")) %>%
  mutate(strip_name = purrr::map(strip_name, purrr::pluck, 1)) %>%
  mutate(strip_name = trimws(paste(strip_name, map))) %>%
  filter(hemisphere == "left") %>%
  select(IDP, strip_name)

# Check for duplicated values in strip name
sum(duplicated(left_hemisphere$strip_name))

```

```{r}

right_hemisphere <- IDP_names %>%
  mutate(strip_name = strsplit(IDP_names$IDP_name, "\\(")) %>%
  mutate(strip_name = purrr::map(strip_name, purrr::pluck, 1)) %>%
  mutate(strip_name = trimws(paste(strip_name, map))) %>%
  filter(hemisphere == "right") %>%
  dplyr::rename("paired" = "IDP") %>%
  select(strip_name, paired)

```

Join the left and right hemispheres over strip_name
Check the length before and after to make sure join doesn't create rows

```{r}
print(paste("Before join:", nrow(left_hemisphere)))
left_hemisphere <- left_join(left_hemisphere, right_hemisphere, by = "strip_name")
print(paste("After join:", nrow(left_hemisphere)))

```
Rename to make columns relevant to the right hemisphere, then stack

```{r}
right_hemisphere <- left_hemisphere %>%
  dplyr::rename("IDP" = "paired", "paired" = "IDP")

paired_cols <- rbind(left_hemisphere, right_hemisphere)

```

Remove strip name (don't need it) and rejoin paired columns onto IDP_names

```{r}
paired_cols <- paired_cols %>%
  select(IDP, paired)

print(paste("Before join:", nrow(IDP_names)))
IDP_names <- left_join(IDP_names, paired_cols, by = "IDP")
print(paste("After join:", nrow(IDP_names)))

```

Create the rfMRI names (these are derived from IDPs)
```{r}

rfMRI_names <- data.frame(matrix(ncol = ncol(IDP_names), nrow = 6))
colnames(rfMRI_names) <- colnames(IDP_names)

rfMRI_names$IDP_name <- c("ICA1", "ICA2", "ICA3", "ICA4", "ICA5", "ICA6")
rfMRI_names$IDP <- c("ICA1", "ICA2", "ICA3", "ICA4", "ICA5", "ICA6")
rfMRI_names$numeric_IDP <- c("1", "2", "3", "4", "5", "6")
rfMRI_names$MRI <- rep("rfMRI", 6)
rfMRI_names$region <- rep("global", 6)


```

Stick the rfMRI names onto the main table
```{r}

IDP_names <- rbind(IDP_names, rfMRI_names)

```


Make a column of numbering for display - want this to be grouped by region

```{r}

IDP_names <- IDP_names %>%
  mutate(display_group = paste0(hemisphere, "_", region)) %>%
  mutate(display_group = gsub("NA_", "", display_group)) # Remove NA that get into the display_group names

```


Save the IDP names table for easy joining to future data sets

```{r}
IDP_name_table <- IDP_names
save(IDP_name_table, file = paste0(output_path, "IDP_names_table_", Sys.Date(), ".RData"))

```

Also save as .csv
```{r}

write.csv(IDP_name_table, file = paste0(output_path, "IDP_names_table_", Sys.Date(), ".csv"))

```


```{r}

print(paste0("Script last run ", Sys.Date()))

```

