---
title: "Smoking covariates"
---


6/27/23 The new phenotypes file from UK Biobank has replaced the phrases such as "All or most days" with a numeric coding system. This step has been substantially re-worked to work with this new system.

An aspect of the smoking data is that some questions were not asked of all participants depending on their responses to earlier questions. In order to determine the rate of missing data and prepare for imputation it is important to determine where data is genuinely missing vs. participants were just not asked.

Overall flow:

ever_daily_smoked: 1 = participants who answered "Yes on most or all days" in Data field 1239 Current tobacco smoking (1 per Data coding 100347) or answered "Smoked on most or all days" (1 per Data coding 100348) in Data field 1249 Past tobacco smoking. 0 = everyone else

never_smoked: 1 = participants who answered "I have never smoked" in Data field 1249 Past tobacco smoking (4 per Data coding 100348) or "No" in Data field 2644 Light smokers, at least 100 smokes in lifetime (0 per Data coding 100349). 0 = everyone else

pack_years: assigned for ever daily smokers by UK Biobank. We further set never smokers to 0 pack years, and light smokers (never daily but also more than 100) to 0.5.


```{r}
library(dplyr) # Run install.packages("tidyverse") to get this
library(tidyr) # Run install.packages("tidyr") to get this
library(ggplot2)
library("reshape") # Run install.packages("reshape") to get this
library("broom")
library(kableExtra) # Run install.packages("kableExtra")

```


Specify to run this on the imaging or baseline smoking responses
Set to "imaging" or "baseline"
Tested this with both options set and it is working 8/1/22
```{r}

touchscreen <- "imaging"

```

Set the filepath to get input files
```{r}
# Home computer
input_path <- "file_path"

```

Set the filepath to save outputs to
```{r}
# Home computer
output_path <- "file_path"

```

Set the date the pipeline is being run
```{r}

run_date <- "yyyy-mm-dd"

```


Load the filtered eid (participants have imaging data)
```{r}

load(paste0(output_path, "filtered_eid_p2_", run_date, ".RData"))
print(paste0(length(filtered_eid)))

```

Pull in the smoking data set (imaging vs. baseline is specified above)
```{r}

smoking_table <- read.csv(paste0(input_path, "smoking_", touchscreen, "_date.csv"))
smoking_table <- filter(smoking_table, n_eid %in% filtered_eid)


```

Strip out the visit indicators on the data field names, so that this script can run on the baseline or imaging data sets
```{r}

names(smoking_table) <- names(smoking_table) %>%
gsub("_0_0", "", .) %>%
gsub("_2_0", "", .)

```

Create the ever daily smoked column
ever_daily_smoked: 1 = participants who answered "Yes on most or all days" in Data field 1239 Current tobacco smoking (1 per Data coding 100347) or answered "Smoked on most or all days" (1 per Data coding 100348) in Data field 1249 Past tobacco smoking. 0 = everyone else
Missing values: NA, -3 "prefer not to answer"
```{r}

smoking_table <- smoking_table %>%
  mutate(ever_daily_smoked = case_when(
    is.na(n_1239) ~ -3,
    n_1239 == -3 ~ -3,
    n_1239 == 1 ~ 1,
    is.na(n_1249) ~ -3,
    n_1249 == -3 ~ -3,
    n_1249 == 1 ~ 1,
    TRUE ~ 0)) %>%
  # Replace the missing values with NA so they can be imputed later
  mutate(ever_daily_smoked = ifelse(ever_daily_smoked == -3, NA, ever_daily_smoked))

smoking_table %>%
  group_by(ever_daily_smoked) %>%
  tally()

```


Create the never smoked column
never_smoked: 1 = participants who answered "I have never smoked" in Data field 1249 Past tobacco smoking (4 per Data coding 100348) or "No" in Data field 2644 Light smokers, at least 100 smokes in lifetime (0 per Data coding 100349). 0 = everyone else

Data field 1249 was not asked for current daily smokers
Data field 2644 was only asked for participants who chose "Smoked occasionally" (2) or "Just tried once or twice" (3) in Data field 1249
```{r}

smoking_table <- smoking_table %>%
  mutate(never_smoked = case_when(
    ever_daily_smoked == 1 ~ 0, # Daily smokers were not asked about past smoking
    is.na(n_1249) ~ -3,
    n_1249 == -3 ~ -3,
    n_1249 == 4 ~ 1,
    # Options 0, 4, and -3 in data field 1249 are dealt with, NA, -1, or -3 in 2644 are true missing data
    is.na(n_2644) ~ -3,
    n_2644 == -1 ~ -3,
    n_2644 == -3 ~ -3,
    n_2644 == 0 ~ 1,
    TRUE ~ 0)) %>%
  # Replace the missing values with NA so they can be imputed later
  mutate(never_smoked = ifelse(never_smoked == -3, NA, never_smoked))

smoking_table %>%
  group_by(never_smoked) %>%
  tally()

```

Add a smoking status column with current, former, never for demographics
Current includes current occasional smokers
Never includes the people who smoked less than 100

1239 Current tobacco smoking, data coding 100347 (https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=100347)
1 = yes on most or all days
2 = only occasionally
0 = No
- 3 = prefer not to answer

1249 Past tobacco smoking, data coding 100348 (https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=100348)
1 = smoked on all or most days
2 = smoked occasionally
3 = just tried once or twice
4 = I have never smoked
-3 = prefer not to answer

2644 Light smokers, at least 100 smokes in lifetime, data coding 100349 (https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=100349)
1 = Yes
0 = No
-1 = Do not know
-3 = Prefer not to answer

```{r}

smoking_table <- smoking_table %>%
  mutate(smoking_status = case_when(
  n_2644 == 0 ~ "Never", # Less than 100 lifetime are Never even if they mention current occasional smoking
  # Daily or occasional current smokers
  n_1239 == 1 ~ "Current",
  n_1239 == 2 ~ "Current",
  # Former daily and true never
  n_1249 == 1 ~ "Former",
  n_1249 == 4 ~ "Never",
  # Former occasional and tried out, get asked 6244
  n_2644 == 1 ~ "Former")) # Everyone else gets NA because they are true unknowns

smoking_table %>%
  group_by(smoking_status) %>%
  tally()

```

Do another check to make sure these categories make sense
```{r}

smoking_table %>%
  group_by(smoking_status, ever_daily_smoked, never_smoked, n_2644) %>%
  tally()

```

pack_years: assigned for ever daily smokers by UK Biobank. We further set never smokers to 0 pack years (never_smokers = 1), and light smokers (never daily but also more than 100) to 0.5 (1 in Data field 2644 per Data coding 100349 is yes to "In your lifetime have you smoked a total of at least 100 times)
```{r}

smoking_table <- smoking_table %>%
  mutate(n_20161 = case_when(
    never_smoked == 1 ~ 0,
    n_2644 == 1 ~ 0.5,
    TRUE ~ n_20161))

```

Get cigarettes per day:
Data field 3456 Number of cigarettes currently smoked daily
Special values from coding 100355
-10 = less than 1 per day (assign 0.1)
-1 = do not know
-3 = prefer not to answer

Data field 2887 Number of cigarettes previously smoked daily
Special values from data coding 100353
-10 = less than 1 per day (assign 0.1)
-1 = do not know

Current occasional smokers and those answering -10 for CPD fields get 0.1
Never and past occasional smokers get 0

```{r}

smoking_table <- smoking_table %>%
  mutate(cig_per_day = case_when(
    # Never smokers have 0 CPD
    smoking_status == "Never" ~ 0,
    # Current occasional smokers assigned 0.1
    n_1239 == 2 ~ 0.1,
    # Past occasional smokers get 0
    n_1249 > 1 ~ 0)) %>%
  # Combine current and former CPD
  mutate(cig_per_day = ifelse(! is.na(n_3456), n_3456, cig_per_day)) %>%
  mutate(cig_per_day = ifelse(! is.na(n_2887), n_2887, cig_per_day)) %>%
  # Handle less than 1 per day from CPD fields
  mutate(cig_per_day = ifelse(cig_per_day == -10, 0.1, cig_per_day)) %>%
  mutate(cig_per_day = ifelse(cig_per_day < 0, NA, cig_per_day))

```



Select the columns that will be used in further processing, and rename so it is easier to read
```{r}

smoking_table <- smoking_table %>%
  dplyr::rename("pack_years" = n_20161) %>%
  select(c(n_eid, pack_years, cig_per_day, ever_daily_smoked, never_smoked, smoking_status))

```

Do a sanity check on pack years values
```{r}

smoking_table %>%
  select(pack_years, never_smoked, ever_daily_smoked) %>%
  group_by(ever_daily_smoked, never_smoked) %>%
  summarise_all(funs(mean, sd, min, max), na.rm = TRUE) %>%
  arrange(desc(ever_daily_smoked)) %>%
  mutate(mean = formatC(mean, format = "f", digits = 2)) %>%
  mutate(sd = formatC(sd, format = "f", digits = 2)) %>%
  kable(caption = paste0("Pack years distribution by smoking status (N=",
                         nrow(smoking_table), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")


```

Sanity check cigarettes per day
```{r}

smoking_table %>%
  select(cig_per_day, smoking_status) %>%
  group_by(smoking_status) %>%
  summarise_all(funs(mean, sd, min, max), na.rm = TRUE) %>%
  mutate(mean = formatC(mean, format = "f", digits = 2)) %>%
  mutate(sd = formatC(sd, format = "f", digits = 2)) %>%
  kable(caption = paste0("CPD distribution by smoking status (N=",
                         nrow(smoking_table), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Some surprizing range values here, per UKB showcase:

The following checks were performed:

If answer < 1 then rejected
If answer > 150 then rejected
If answer > 100 then participant asked to confirm

So the outlier participant/s confirmed this input

Save the processed smoking variables, the name will automatically adjust if it is the imaging or baseline set
Also save the missing report
```{r}

if (touchscreen == "baseline") {
  baseline_smoking <- smoking_table
  save(baseline_smoking, file = paste0(output_path, "baseline_smoking_", Sys.Date(), ".RData"))
}

if (touchscreen == "imaging") {
  imaging_smoking <- smoking_table
  save(imaging_smoking, file = paste0(output_path, "imaging_smoking_", Sys.Date(), ".RData"))
}

```


```{r}

print(paste0("Script last run ", Sys.Date()))

```



