---
title: "Pipeline Step 4 - Alcohol"
output:
  html_document:
    df_print: paged
---

Process the alcohol variables
Add anything missing to the missing report
```{r}
library(dplyr) # Run install.packages("tidyverse") to get this
library(tidyr) # Run install.packages("tidyr") to get this
library(ggplot2)
library("reshape") # Run install.packages("reshape") to get this
library("broom")
library(kableExtra) # Run install.packages("kableExtra")

```


Set the filepath to get input files
```{r}
# Home computer
input_path <- "file_path"

```

Set the filepath to save outputs to
```{r}
# Home computer
output_path <- "file_path"

```

Set the date the IDP names table script was run
```{r}

run_date <- "yyyy-mm-dd"

```


Load the filtered eid and missing_report
```{r}

load(paste0(output_path, "filtered_eid_p2_", run_date, ".RData"))
load(paste0(output_path, "missing_report_p2_", run_date, ".RData"))

```

Load the imaging and baseline alcohol files
```{r}

baseline <- read.csv(paste0(input_path, "alcohol_baseline.csv"))
baseline <- filter(baseline, n_eid %in% filtered_eid)
names(baseline) <- names(baseline) %>%
gsub("_0_0", "", .) %>%
gsub("_2_0", "", .)

imaging <- read.csv(paste0(input_path, "alcohol_imaging.csv"))
imaging <- filter(imaging, n_eid %in% filtered_eid)
names(imaging) <- names(imaging) %>%
gsub("_0_0", "", .) %>%
gsub("_2_0", "", .)

```


Calculate number of drinks per week and drinks per drinking day
UK Bio has the number of drinks split by drink type (ex. red wine, beer)
3/14/22 Used to have month dose as the output, fixed this to week dose

The data fields representing estimated number of drinks use data coding 100291
Valid responses are positive numbers representing drinks
-1 = Do not know
-3 = Prefer not to answer
https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=100291

```{r}

imaging %>%
  # Make a column to indicate monthly or weekly frequency
  mutate(week_month = case_when(
    is.na(n_1558) ~ "missing",
    n_1558 == -3 ~ "missing",
    n_1558 <= 3 ~ "week",
    n_1558 < 6 ~ "month",
    n_1558 == 6 ~ "never",
    TRUE ~ "missing")) %>%
  mutate(n_drinks = NA) %>%
  
  # Convert any strings to numeric
  mutate_at(vars(n_1568, n_1578, n_1588, n_1598, n_1608, n_5364), ~as.numeric(.x)) %>%
  
  # Change any special (-1, -3) values to NA
  mutate_at(vars(n_1568, n_1578, n_1588, n_1598, n_1608, n_5364), ~ifelse(.x < 0, NA, .x)) %>%
  
  filter(n_1568 < 0)


```




```{r}

imaging <- imaging %>%
  # Make a column to indicate monthly or weekly frequency
  mutate(week_month = case_when(
    is.na(n_1558) ~ "missing",
    n_1558 == -3 ~ "missing",
    n_1558 <= 3 ~ "week",
    n_1558 < 6 ~ "month",
    n_1558 == 6 ~ "never",
    TRUE ~ "missing")) %>%
  mutate(n_drinks = NA) %>%
  
  # MONTHLY
  # Convert any strings to numeric
  mutate_at(vars(n_4407, n_4418, n_4429, n_4440, n_4451, n_4462), ~as.numeric(.x)) %>%
  
  # Change any special (-1, -3) values to NA
  mutate_at(vars(n_4407, n_4418, n_4429, n_4440, n_4451, n_4462), ~ifelse(.x < 0, NA, .x)) %>%
  
  rowwise() %>%
  # Add all the different drink types
  mutate(n_drinks = ifelse(week_month == "month",
    sum(n_4407, n_4418, n_4429, n_4440, n_4451, n_4462, na.rm = TRUE), n_drinks)) %>%
  ungroup() %>%
  # If all categories were NA, put NA back as the sum (it gets turned to 0 by na.rm = TRUE)
  mutate(n_drinks = ifelse(week_month == "month",
                           ifelse(is.na(n_4407) & is.na(n_4418) & is.na(n_4429) & is.na(n_4440)
                             & is.na(n_4451) & is.na(n_4462), NA, n_drinks), n_drinks)) %>%
  
  # WEEKLY
  # Convert any strings to numeric
  mutate_at(vars(n_1568, n_1578, n_1588, n_1598, n_1608, n_5364), ~as.numeric(.x)) %>%
  
  # Change any special (-1, -3) values to NA
  mutate_at(vars(n_1568, n_1578, n_1588, n_1598, n_1608, n_5364), ~ifelse(.x < 0, NA, .x)) %>%
  
  rowwise() %>%
  mutate(n_drinks = ifelse(week_month == "week",
    sum(n_1568, n_1578, n_1588, n_1598, n_1608, n_5364, na.rm = TRUE), n_drinks)) %>%
  ungroup() %>%
  # If all categories were NA, put NA back as the sum (it gets turned to 0 by na.rm = TRUE)
  mutate(n_drinks = ifelse(week_month == "week",
                           ifelse(is.na(n_1568) & is.na(n_1578) & is.na(n_1588) & is.na(n_1598)
                             & is.na(n_1608) & is.na(n_5364), NA, n_drinks), n_drinks)) %>%
  
  # Assign never drinkers to 0 drinks
  mutate(n_drinks = ifelse(week_month == "never", 0, n_drinks)) %>%
  
  # Convert n_drinks into drinks per week
  mutate(week_drinks = ifelse(week_month == "month", n_drinks / 4.34, n_drinks)) %>% # 4.34 weeks per month
  
  # Convert data field 1558 into drinking days per month
  mutate(drink_freq = case_when(
    n_1558 == 1 ~ 30.4, # Average number of days in a month
    n_1558 == 2 ~ 3.5 * 4.3, # Average days per week * average # weeks in a month
    n_1558 == 3 ~ 1.5 * 4.3, # Average days per week * average # weeks in a month
    n_1558 == 4 ~ 2, # Average times per month
    n_1558 == 5 ~ 0.5, # Less than monthly
    n_1558 == 6 ~ 0))
  

```

Repeat on the baseline alcohol data

```{r}

baseline <- baseline %>%
  # Make a column to indicate monthly or weekly frequency
  mutate(week_month = case_when(
    is.na(n_1558) ~ "missing",
    n_1558 == -3 ~ "missing",
    n_1558 <= 3 ~ "week",
    n_1558 < 6 ~ "month",
    n_1558 == 6 ~ "never",
    TRUE ~ "missing")) %>%
  mutate(n_drinks = NA) %>%
  
  # Monthly
  # Convert any strings to numeric
  mutate_at(vars(n_4407, n_4418, n_4429, n_4440, n_4451, n_4462), ~as.numeric(.x)) %>%
  
  # Change any special (-1, -3) values to NA
  mutate_at(vars(n_4407, n_4418, n_4429, n_4440, n_4451, n_4462), ~ifelse(.x < 0, NA, .x)) %>%
  
  rowwise() %>%
  # Add all the different drink types
  mutate(n_drinks = ifelse(week_month == "month",
    sum(n_4407, n_4418, n_4429, n_4440, n_4451, n_4462, na.rm = TRUE), n_drinks)) %>%
  ungroup() %>%
  # If all categories were NA, put NA back as the sum (it gets turned to 0 by na.rm = TRUE)
  mutate(n_drinks = ifelse(week_month == "month",
                           ifelse(is.na(n_4407) & is.na(n_4418) & is.na(n_4429) & is.na(n_4440)
                             & is.na(n_4451) & is.na(n_4462), NA, n_drinks), n_drinks)) %>%
  
  # WEEKLY
  # Convert any strings to numeric
  mutate_at(vars(n_1568, n_1578, n_1588, n_1598, n_1608, n_5364), ~as.numeric(.x)) %>%
  
  # Change any special (-1, -3) values to NA
  mutate_at(vars(n_1568, n_1578, n_1588, n_1598, n_1608, n_5364), ~ifelse(.x < 0, NA, .x)) %>%

  rowwise() %>%
  mutate(n_drinks = ifelse(week_month == "week",
    sum(n_1568, n_1578, n_1588, n_1598, n_1608, n_5364, na.rm = TRUE), n_drinks)) %>%
  ungroup() %>%
  # If all categories were NA, put NA back as the sum (it gets turned to 0 by na.rm = TRUE)
  mutate(n_drinks = ifelse(week_month == "week",
                           ifelse(is.na(n_1568) & is.na(n_1578) & is.na(n_1588) & is.na(n_1598)
                             & is.na(n_1608) & is.na(n_5364), NA, n_drinks), n_drinks)) %>%
  
  # Assign never drinkers to 0 drinks
  mutate(n_drinks = ifelse(week_month == "never", 0, n_drinks)) %>%
  
  # Convert n_drinks into drinks per week
  mutate(week_drinks = ifelse(week_month == "month", n_drinks / 4.34, n_drinks)) %>% # 4.34 weeks per month
  
  # Convert data field 1558 into drinking days per month
  mutate(drink_freq = case_when(
    n_1558 == 1 ~ 30.4, # Average number of days in a month
    n_1558 == 2 ~ 3.5 * 4.3, # Average days per week * average # weeks in a month
    n_1558 == 3 ~ 1.5 * 4.3, # Average days per week * average # weeks in a month
    n_1558 == 4 ~ 2, # Average times per month
    n_1558 == 5 ~ 0.5, # Less than monthly
    n_1558 == 6 ~ 0))
  

```


Make a column to reflect people who are true never drinkers vs. former drinkers
Data Field 3731 Former alcohol drinker (UKB doesn't always use person first language)
1 = Yes
0 = No
-3 = Prefer not to answer
```{r}

imaging <- imaging %>%
  mutate(never_alcohol = case_when(
    week_month == "week" ~ 0,
    week_month == "month" ~ 0,
    week_month == "never" ~ ifelse(n_3731 == 0, 1, 0))) # True never consumed alcohol
    # "missing" from 1558 will be NA
    # -3 from 3731 is False for never consumed alcohol, we just don't know so err on conservative side

baseline <- baseline %>%
  mutate(never_alcohol = case_when(
    week_month == "week" ~ 0,
    week_month == "month" ~ 0,
    week_month == "never" ~ ifelse(n_3731 == 0, 1, 0))) # True never consumed alcohol
    # "missing" from 1558 will be NA


```

Within no-drinking, also capture how many are current, former, or truly never
Data-Field 3731, Former alcohol drinker, asked only of those who indicated "Never" in data field 1558

Data-Coding 100352
1 = Yes
2 = No
-3 = Prefer not to answer

```{r}

imaging <- imaging %>%
  mutate(alcohol_status = NA) %>%
  mutate(alcohol_status = case_when(
    week_month == "week" ~ "Current",
    week_month == "month" ~ "Current",
    week_month == "never" ~ ifelse(n_3731 == 0, "Never", "Former")))
    # "missing" from 1558 will be NA
    # -3 from 3731 is False for never consumed alcohol, we just don't know so err on conservative side

baseline <- baseline %>%
  mutate(alcohol_status = NA) %>%
  mutate(alcohol_status = case_when(
    week_month == "week" ~ "Current",
    week_month == "month" ~ "Current",
    week_month == "never" ~ ifelse(n_3731 == 0, "Never", "Former")))
    # "missing" from 1558 will be NA
    # -3 from 3731 is False for never consfumed alcohol, we just don't know so err on conservative side


```


Check that the pattern of drink freq, non-drinker, and alcohol status make sense
```{r}

imaging %>%
  group_by(drink_freq, never_alcohol, alcohol_status) %>%
  tally()

```

Set up a missing report to capture how much data is missing before and after replacement
```{r}

missing_report <- rbind(missing_report, data.frame("step" = rep("Before"),
                             "data_field" = c("drink_frequency", "week_drinks", "never_alcohol", "alcohol_status"),
                             "n_missing" = c(sum(is.na(imaging$drink_freq)),
                                             sum(is.na(imaging$week_drinks)),
                                             sum(is.na(imaging$never_alcohol)),
                                             sum(is.na(imaging$alcohol_status)))))

```


Where the imaging visit alcohol data is missing values, replace with the baseline visit

```{r}

baseline <- baseline %>%
  dplyr::rename("baseline_drink_freq" = drink_freq,
                "baseline_week_drinks" = week_drinks,
                "baseline_never_alcohol" = never_alcohol,
                "baseline_alcohol_status" = alcohol_status)

imaging <- imaging %>%
  left_join(baseline, by = "n_eid") %>%
  mutate(drink_freq = ifelse(is.na(drink_freq), baseline_drink_freq, drink_freq)) %>%
  mutate(week_drinks = ifelse(is.na(week_drinks), baseline_week_drinks, week_drinks)) %>%
  mutate(never_alcohol = ifelse(is.na(never_alcohol), baseline_never_alcohol, never_alcohol)) %>%
  mutate(alcohol_status = ifelse(is.na(alcohol_status), baseline_alcohol_status, alcohol_status))

```

Capture the number missing after the replacement step

```{r}

missing_report <- rbind(missing_report, data.frame("step" = rep("After"),
                             "data_field" = c("drink_frequency", "week_drinks", "never_alcohol", "alcohol_status"),
                             "n_missing" = c(sum(is.na(imaging$drink_freq)),
                                             sum(is.na(imaging$week_drinks)),
                                             sum(is.na(imaging$never_alcohol)),
                                             sum(is.na(imaging$alcohol_status)))))

```


```{r}

missing_report %>%
  mutate(percent_missing = n_missing / nrow(imaging) * 100) %>%
  mutate(percent_missing = formatC(percent_missing, digits = 1, format = "f")) %>%
  kable(caption = paste0("Missing values in Imaging Visit Alcohol Data Before and After Replacement from Baseline (N=",
                         nrow(imaging), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```


Rename and select to just the essential alcohol columns
```{r}

imaging_alcohol <- imaging %>%
  select(n_eid, drink_freq, week_drinks, never_alcohol, alcohol_status)

rm(baseline)
rm(imaging)

```

Check range and distribution to ensure it makes sense
```{r}

imaging_alcohol %>%
  select(c(week_drinks, drink_freq)) %>%
  summarise_all(funs(mean, sd, min, max), na.rm = TRUE) %>%
  pivot_longer(cols = everything(), names_to = "name", values_to = "value") %>%
  mutate(measure = case_when(
    grepl("_mean", name) ~ "mean",
    grepl("_sd", name) ~ "sd",
    grepl("_min", name) ~ "min",
    grepl("_max", name) ~ "max")) %>%
  mutate(name = gsub("_mean", "", name)) %>%
  mutate(name = gsub("_sd", "", name)) %>%
  mutate(name = gsub("_min", "", name)) %>%
  mutate(name = gsub("_max", "", name)) %>%
  pivot_wider(id_cols = name, names_from = measure, values_from = value) %>%
  mutate(mean = formatC(mean, digits = 2, format = "f")) %>%
  mutate(sd = formatC(sd, digits = 2, format = "f"))  %>%
  mutate(min = formatC(min, digits = 2, format = "f"))  %>%
  mutate(max = formatC(max, digits = 2, format = "f"))  %>%
  kable(caption = paste0("Mean, SD, Min, and Max in Alcohol Covariates (N=",
                         nrow(imaging_alcohol), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")
  
  
```

Save the alcohol variables and missing report
```{r}

save(imaging_alcohol, file = paste0(output_path, "imaging_alcohol_", Sys.Date(), ".RData"))
save(missing_report, file = paste0(output_path, "missing_report_p4_", Sys.Date(), ".RData"))

```


```{r}

print(paste0("Script last run ", Sys.Date()))

```


