---
title: "Pipeline Step 1 - Filter participants and import imaging data"
output:
  html_document:
    df_print: paged
---

In this script the imaging data (as IDPs) is imported and participants are filtered for inclusion in the analysis.
Participants are excluded for:
Not being matched to a pheno ID
Indicated they want to withdraw from a study
Have a neurological diagnosis which could affect their imaging
Are a 3rd degree or closer relative of someone else in the study (which one out of a relative pair is retained is random)

Not all participants have all MRI sequences, so the IDPs are split by sequence and participants are dropped from the set if they are missing the sequence.

```{r}
library(dplyr) # Run install.packages("tidyverse") to get this
library(tidyr) # Run install.packages("tidyr") to get this
library(ggplot2)
library("reshape") # Run install.packages("reshape") to get this
library(kableExtra) # Run install.packages("kableExtra")
library(viridis)
library(forcats)

```

Set the filepath to get input files
```{r}
# Home computer
input_path <- "file_path"

```

Set the filepath to save outputs to
```{r}
# Home computer
output_path <- "file_path"

```

Set the date the IDP names table script was run
```{r}

run_date <- "yyyy-mm-dd"

```


Initialize a table to track N as the data set is processed
```{r}

n_tracker <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(n_tracker) <- c("Step", "MRI", "N")

```

Import the file of IDPs

```{r}

ukbio_IDP <- read.csv(file = paste0(input_path, "my_UKB_IDP_data.csv"))

```

Capture n in the full imaging set
```{r}

n_tracker <- rbind(n_tracker, data.frame("Step" = "Imaging cohort", "MRI" = "All", "N" = nrow(ukbio_IDP)))

```

Stitch in the rfMRI ICAs (also have the imaging ID)
```{r}

rfMRI_ICA <- read.table(file = paste0(input_path, "subs_netica.txt"), sep = ",", header = FALSE)

# Do some renaming, the first column is the subject IDs
rfMRI_ICA <- rfMRI_ICA %>%
  dplyr::rename("eid" = V1, "ICA1" = V2, "ICA2" = V3, "ICA3" = V4, "ICA4" = V5, "ICA5" = V6, "ICA6" = V7)

```

```{r}

ukbio_IDP <- left_join(ukbio_IDP, rfMRI_ICA, by = "eid")

```


Replace the imaging eid with the phenotype eid so they will match all the way through

The column entitled ukb47267_ID_1 contains the IDs present in the imaging set.
The column ukb48123_ID_1 is the IDs in the phenotype set.

```{r}

ID_key <- read.table(file = paste0(input_path, "/ID_key_for_my_UKB.txt"), sep = " ", header = TRUE)

ID_key <- dplyr::rename(ID_key, pheno_eid = ukb48123_ID_1, eid = ukb47267_ID_1)

```

Now join these columns into the imaging data over ukb47267_ID_1 (eid)
```{r}
ukbio_IDP <- right_join(ID_key, ukbio_IDP, by = "eid")

```


Filter out the patients who are missing an eid - we cannot match them to the phenotypes
```{r}

ukbio_IDP <- ukbio_IDP %>%
  filter(! is.na(pheno_eid))

```


Update the n tracker
```{r}

n_tracker <- rbind(n_tracker, data.frame("Step" = "Matched phenotype eid", "MRI" = "All", "N" = nrow(ukbio_IDP)))

```

Remove the imaging eid (no longer needed) and rename the pheno_eid so it will match the phenotyping and covariates

```{r}

ukbio_IDP <- ukbio_IDP %>%
  select(! eid) %>%
  dplyr::rename("n_eid" = "pheno_eid")

```

Filter out withdrawn participants
```{r}

withdrawn_ID <- read.csv(paste0(input_path, "UKB_withdrawn.csv"))
ukbio_IDP <- ukbio_IDP %>%
  filter(! ukbio_IDP$n_eid %in% withdrawn_ID$X1038692)

```

Save the n
```{r}

n_tracker <- rbind(n_tracker, data.frame("Step" = "Withdrawn removed", "MRI" = "All", "N" = nrow(ukbio_IDP)))

```


================Remove participants with neurological disease===========================================

4/29/22 Had been using strategy of removing all participants with any IDP an outlier by >4 SD
Checked this against known neuro disease, it only catches 30% of participants with known disease
While also removing 30% of the sample
Decided to simply use the diagnosis codes from data field 20002 to filter out participants and leave the outliers alone

Assess how many participants are >4 SD outliers, and how many have a neurological disease diagnosis

6/15/23 Include the disease list from both baseline and imaging appointments, they are different
Participants filtered out if the disease is reported at any appointment

```{r}

disease_baseline <- read.csv(paste0(input_path, "disease_baseline_6_13_23.csv"))
disease_imaging <- read.csv(paste0(input_path, "disease_imaging_6_13_23.csv"))

```

These data fields include all disease codes, although we are using them to filter for neurological disease only

Combine baseline and imaging, filter for only the people in our imaging cohort
```{r}

all_disease <- disease_baseline %>%
  left_join(disease_imaging, by = "n_eid") %>%
  filter(n_eid %in% ukbio_IDP$n_eid)

```


Make a list of codes for neurological diseases that might impact brain structure
I referenced the Palmer paper, will need to verify against UK Biobank and also confirm we agree with this list
Dementia = 1263
Parkinsons = 1262
Chronic degenerative neurological = 1258
Guillan-Barre syndrome = 1256
Multiple sclerosis = 1261
Other demyelinating disease = 1397
Stroke or ischaemic stroke = 1081
Brain cancer = 1032
Brain hemorrhage = 1491
Brain / intracranial abcess = 1245
Cerebral aneurysm = 1425
Cerebral palsy = 1433
Encephalitis = 1246
Epilepsy = 1264
Head injury = 1266
Nervous system infection = 1244
Ischaemic stroke = 1583
Meningeal cancer = 1031
Meningioma = 1659
Meningitis = 1247
Motor neuron disease = 1259
Neurological disease / trauma = 1240
Spina bifida = 1524
Subdural hematoma = 1083
Subarachnoid hemorrhage = 1086
Transient ischemic attack = 1082


```{r}

neuro_codes <- data.frame((rbind(
  c("Dementia", 1263),
  c("Parkinsons", 1262),
  c("Chronic degenerative neurological", 1258),
  c("Guillan-Barre syndrome", 1256),
  c("Multiple sclerosis", 1261),
  c("Other demyelinating disease", 1397),
  c("Stroke or ischaemic stroke", 1081),
  c("Brain cancer", 1032),
  c("Brain hemorrhage", 1491),
  c("Brain / intracranial abscess", 1245),
  c("Cerebral aneurysm", 1425),
  c("Cerebral palsy", 1433),
  c("Encephalitis", 1246),
  c("Epilepsy", 1264),
  c("Head injury", 1266),
  c("Nervous system infection", 1244),
  c("Ischaemic stroke", 1583),
  c("Meningeal cancer", 1031),
  c("Meningioma", 1659),
  c("Meningitis", 1247),
  c("Motor neuron disease", 1259),
  c("Neurological disease / trauma", 1240),
  c("Spina bifida", 1524),
  c("Subdural hematoma", 1083),
  c("Subarachnoid hemorrhage", 1086),
  c("Transient ischemic attack", 1082))))

neuro_codes <- neuro_codes %>%
  dplyr::rename("condition" = X1, "code" = X2) %>%
  mutate(code = formatC(code, format = "d"))

```

Use the table of neuro codes to construct a table for neuro disease
```{r}

neuro_disease <- all_disease %>%
  pivot_longer(cols = -n_eid, names_to = "field", values_to = "code") %>%
  select(-field) %>%
  filter(! is.na(code)) %>%
  # Remove any duplicates due to data from both the baseline and imaging visits
  distinct(.keep_all = TRUE) %>%
  mutate(code = formatC(code, format = "d")) %>%
  filter(code %in% neuro_codes$code) %>%
  left_join(neuro_codes, by = "code")

```

Use the table to tally up how many total individuals are affected, and which diseases

```{r}

neuro_disease_table <- neuro_disease %>%
  group_by(condition) %>%
  tally() %>%
  left_join(neuro_codes, by = "condition") %>%
  relocate(code, .before = n) %>%
  mutate(percent = n / nrow(ukbio_IDP) * 100) %>%
  mutate(percent = formatC(percent, digits = 2, format = "f")) %>%
  arrange(desc(n))

neuro_disease_table %>%
  kable(caption = paste0("Prevalence of neurological conditions in Imaging cohort\nN=",
                         nrow(ukbio_IDP), ", N affected=", length(unique(neuro_disease$n_eid)))) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Save the neuro disease table, it will go in the supplemental info
```{r}

save(neuro_disease_table, file = paste0(output_path, "neuro_disease_table_", Sys.Date(), ".RData"))

```


Use the eids in the neuro_disease table to filter out participants with a diagnosis of neurological disease

```{r}

ukbio_IDP <- ukbio_IDP %>%
  filter(! n_eid %in% neuro_disease$n_eid)

```

Update the N tracker

```{r}

n_tracker <- rbind(n_tracker, data.frame("Step" = "Neuro diagnosis removed", "MRI" = "All", "N" = nrow(ukbio_IDP)))

```

Filter for relatedness=============================================================================

Import the relatedness matrix
Got this from the database - it was uploaded by Louis
```{r}

relatedness <- read.delim(file = paste0(input_path, "ukb48123_rel_s488285.dat"), header = TRUE, sep=" ")

```

Per Bycroft 2018 Supplemental Info, 1st degree relatives have an expected kinship coefficient of 1/4
2nd degree = 1/8
3rd degree = 1/16
Set the minimum kinship coefficient to 1/16th and then filter the relationship matrix for 1/6 or greater
This is a stochastic probability so I set it slightly under 1/16th to capture these

The ukbtools function to accomplish this sets a larger cutoff for related-ness

```{r}

kinship_coeff = 1/20

relatedness <- relatedness %>%
  filter(Kinship > kinship_coeff)

```

This reduces the number of entries in the kinship matrix down to 92825

Filter again for eid in the imaging set
We only care about pairs where both are in the imaging data
When one member of a pair did not, happy day no conflict!

```{r}

relatedness <- relatedness %>%
  filter(relatedness$ID1 %in% ukbio_IDP$n_eid)

relatedness <- relatedness %>%
  filter(relatedness$ID2 %in% ukbio_IDP$n_eid)

```

886 pairs of 3rd degree or closer relatives in the imaging set

For selecting which members to retain, I am implementing the algorithm used in the ukbiotools package by Ken Hanscombe
Description of the algorithm:
1) Remove pairs with kinship coefficient less than 3rd degree (already done)
2) Count number of connections each participant has, add the one with the most connections to "samples to exclude"
  - accomplish this with a while loop
3) Repeat step 2 until all participants have only 1 connection, then randomly select one member of the pairs to add to "samples to exclude"


```{r}

keep_list = c()
remove_list = c()

done = FALSE
while (! done) {
remove_ID <- relatedness %>%
  select(ID1, ID2) %>%
  pivot_longer(everything(), names_to = "colname", values_to = "ID") %>%
  group_by(ID) %>%
  tally() %>%
  arrange(desc(n)) %>%
  utils::head(n = 1)

if (remove_ID$n > 1) {
  # Add the highly connected ID to the remove list
  remove_list <- append(remove_list, remove_ID$ID)
  
  # Filter the highly connected ID out of the relationship matrix
  relatedness <- relatedness %>%
    filter(! (ID1 == remove_ID$ID)) %>%
    filter(! (ID2 == remove_ID$ID))
}  else {
  done = TRUE
}
}

```

All highly connected eIDs are now removed.
Randomly assign the remaining IDs to remove
```{r}

for (i in 1:nrow(relatedness)) {
  if (i %% 2 == 1) {
    remove_list <- append(remove_list, relatedness[i,]$ID1)
  } else
    remove_list <- append(remove_list, relatedness[i,]$ID2)
}

```

Use the remove list to filter the imaging set

```{r}

ukbio_IDP <- ukbio_IDP %>%
  filter(! ukbio_IDP$n_eid %in% remove_list)

```

Save the n after filtering
```{r}

n_tracker <- rbind(n_tracker, data.frame("Step" = "Related removed", "MRI" = "All", "N" = nrow(ukbio_IDP)))

```


======Handle missing data in the imaging IDPs and normalize================================================

Import the IDP names file to allow exploring the IDPs by MRI type, etc.
```{r}

load(paste0(output_path, "IDP_names_table_", run_date, ".RData"))

```

Grab the ukbio_IDP column names to get a list of the IDPs present in the data set
```{r}

IDP_all_names <- colnames(ukbio_IDP)
IDP_all_names <- IDP_all_names[IDP_all_names != "n_eid"]

```

Filter the names table to match the IDPs in our UK Bio imaging data set
4/28/22 modified pipeline step 0 to include the ICAs
```{r}

IDP_name_table <- IDP_name_table %>%
  filter(IDP_name_table$IDP %in% IDP_all_names)

```


Remove the QC IDPs, we are not analyzing these
```{r}

IDP_name_table <- IDP_name_table %>%
  filter(! MRI == "QC") %>%
  # MRI == "resting_functional" are QC related to rfMRI
  filter(! MRI == "resting_functional")

```


Save the filtered list as IDP_all_names
```{r}

IDP_all_names <- IDP_name_table %>%
  pull(IDP)

```

Use IDP_all_names to remove the L1, L2, and L2 dMRI measures from the main imaging set
Also removing tfMRI (task functional)
```{r}

ukbio_IDP <- ukbio_IDP %>%
  select(c(n_eid, all_of(IDP_all_names)))

```

6/16/23 there are 1903 IDPs


Create a table to display the number missing and how that compares as a percentage of the total data
```{r}

image_missing <- ukbio_IDP %>%
  summarise_at(vars(all_of(IDP_all_names)), ~ sum(is.na(.x))) %>%
  pivot_longer(cols = everything(), names_to = "IDP", values_to = "n_missing") %>%
  mutate(percent_missing = n_missing / nrow(ukbio_IDP) * 100) %>%
  left_join(IDP_name_table, by = "IDP")
  
```

Plot the percent missing by MRI type

```{r}

image_missing %>%
  mutate(MRI = ifelse(MRI == "T1", ifelse(map == "global" | map == "FIRST", "T1 global / FIRST",
                                          ifelse(map == "FAST", "T1 FAST", "T1 regional")), MRI)) %>%
  mutate(MRI = fct_reorder(MRI, -percent_missing)) %>%
  ggplot(aes(x = reorder(IDP, -percent_missing), y = percent_missing, color = MRI)) +
  geom_point() +
  ggtitle(paste0("Missing values per IDP as Percent of N (N=", nrow(ukbio_IDP), ")")) +
  xlab(paste0("IDP (IDPs =", nrow(image_missing), ")")) +
  ylab("Percent of observations") +
  ylim(c(0, 8))

```

Group by MRI type and display % missing

```{r}

image_missing %>%
  mutate(MRI = ifelse(IDP == "X25781.2.0", "total_hyperintensities", MRI)) %>%
  group_by(MRI, map) %>%
  summarise(percent_missing = mean(percent_missing)) %>%
  arrange(desc(percent_missing)) %>%
  mutate(percent_missing = formatC(percent_missing, digits = 3, format = "f")) %>%
  kable(caption = paste0("Missing values per MRI Category as Percent of N (N=", nrow(ukbio_IDP), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")
  

```


Participants who are missing all values for a given MRI category should not be included in normalization / regression for that category
Break the imaging data up based on category

Some participants are missing brain volume, they cannot be included in any subsets
```{r}

ukbio_IDP <- ukbio_IDP %>%
  filter(! is.na(X25009.2.0))

```

Break the rest of the IDPs into subsets
Define the total volume IDPs (also used for controlling reginal IDPs) so they do not get included in regional subsets (this causes errors)
25001	Volume of peripheral cortical grey matter (normalised for head size)	T1 structural brain MRI  
25003	Volume of ventricular cerebrospinal fluid (normalised for head size)	T1 structural brain MRI  
25005	Volume of grey matter (normalised for head size)	T1 structural brain MRI  
25007	Volume of white matter (normalised for head size)	T1 structural brain MRI  
25009	Volume of brain, grey+white matter (normalised for head size)	T1 structural brain MRI  
26514	Volume of BrainSeg (whole brain)	Freesurfer ASEG  
26518	Volume of TotalGray (whole brain)	Freesurfer ASEG  
26527	Volume of CSF (whole brain)	Freesurfer ASEG  
26528	Volume of WM-hypointensities (whole brain)	Freesurfer ASEG  
26553	Volume of CerebralWhiteMatter (left hemisphere)	Freesurfer ASEG  
26584	Volume of CerebralWhiteMatter (right hemisphere)	Freesurfer ASEG  
26755	Mean thickness of GlobalMeanMean thickness (left hemisphere)	Freesurfer desikan white  
26856	Mean thickness of GlobalMeanMean thickness (right hemisphere)	Freesurfer desikan white

Add additional total measures from ASEG 11/10/23
26517 Volume of SubCortGray (whole brain) - subcortical grey matter
26583 Volume of Cortex (right hemisphere)
26552 Volume of Cortex (left hemisphere)

Also including IDP derived measures
```{r}

total_volume_IDPs <- c("X25001.2.0", "X25003.2.0", "X25005.2.0", "X25007.2.0", "X25009.2.0",
                       "X26514.2.0", "X26518.2.0", "X26527.2.0", "X26528.2.0", "X26553.2.0", "X26584.2.0",
                       "X26755.2.0", "X26856.2.0", "X26517.2.0", "X26583.2.0", "X26552.2.0", "total_wm",
                       "global_mean_thk", "total_cortex",
                       "mean_FA", "mean_MD", "mean_ICVF", "mean_ISOVF", "mean_MO", "mean_OD")

```


```{r}

GMV <- IDP_name_table %>%
  filter(MRI == "T1") %>%
  filter(map == "global" | map == "FAST" | map == "FIRST") %>%
  filter(! IDP %in% total_volume_IDPs) %>%
  pull(IDP)

length(GMV)

```

```{r}

free_2009 <- IDP_name_table %>%
  filter(MRI == "T1") %>%
  filter(map == "freesurfer_a2009s") %>%
  filter(! IDP %in% total_volume_IDPs) %>%
  pull(IDP)

length(free_2009)

```


```{r}

desikan <- IDP_name_table %>%
  filter(MRI == "T1") %>%
  filter(grepl("desikan", map)) %>%
  filter(! IDP %in% total_volume_IDPs) %>%
  pull(IDP)

length(desikan)

```


```{r}

T1_regional <- IDP_name_table %>%
  filter(MRI == "T1") %>%
  filter(! map %in% c("global", "FAST", "FIRST", "freesurfer_a2009s")) %>%
  filter(! grepl("desikan", map)) %>%
  filter(! IDP %in% total_volume_IDPs) %>%
  pull(IDP)

length(T1_regional)

```

Select T2 with the exception of total hyperintensities (X25781.2.0)

```{r}

T2 <- IDP_name_table %>%
  filter(MRI == "T2") %>%
  filter(! IDP == "X25781.2.0") %>%
  filter(! IDP %in% total_volume_IDPs) %>%
  pull(IDP)

length(T2)

```

Include total hyperintensities with the dMRI measures
It has a similar rate of missingness, so this makes sense
Palmer 2021 groups it in this way

```{r}

dMRI <- IDP_name_table %>%
  filter(MRI == "dMRI") %>%
  filter(! IDP %in% total_volume_IDPs) %>%
  pull(IDP)
  #append("X25781.2.0") # Not in version from Spring 2023

length(dMRI)

```

```{r}

rfMRI <- IDP_name_table %>%
  filter(MRI == "rfMRI") %>%
  pull(IDP)

length(rfMRI)

```

Check that the number of IDPs in the individual vectors matches the total

```{r}

print(sum(length(total_volume_IDPs), length(GMV), length(free_2009), length(desikan), length(T1_regional), length(T2), length(dMRI), length(rfMRI)))

```
Now break the imaging set into subsets based on the IDPs

```{r}

GMV <- ukbio_IDP %>%
  select(c(n_eid, all_of(GMV)))

free_2009 <- ukbio_IDP %>%
  select(c(n_eid, all_of(free_2009)))

desikan <- ukbio_IDP %>%
  select(c(n_eid, all_of(desikan)))

T1_regional <- ukbio_IDP %>%
  select(c(n_eid, all_of(T1_regional)))

T2 <- ukbio_IDP %>%
  select(c(n_eid, all_of(T2)))

dMRI <- ukbio_IDP %>%
  select(c(n_eid, all_of(dMRI)))

rfMRI <- ukbio_IDP %>%
  select(c(n_eid, all_of(rfMRI)))

```

Filter each subset to remove participants who are missing data

```{r}

GMV <- na.omit(GMV)
free_2009 <- na.omit(free_2009)
desikan <- na.omit(desikan)
T1_regional <- na.omit(T1_regional)
T2 <- na.omit(T2)
dMRI <- na.omit(dMRI)
rfMRI <- na.omit(rfMRI)

```

Add rows to the N tracker

```{r}

MRI_types <- c("GMV", "free_2009", "desikan", "T1_regional", "T2", "dMRI", "rfMRI")

n_tracker <- rbind(n_tracker, data.frame("Step" = rep("MRI missing dropped", length(MRI_types)),
            "MRI" = MRI_types,
           "N" = c(nrow(GMV), nrow(free_2009), nrow(desikan), nrow(T1_regional),
                   nrow(T2), nrow(dMRI), nrow(rfMRI))))

```

############################## Calculate additional total measures and control measures

Start a table for IDPs and IDP derived measures that are used as controls in the regression analysis
Control columns:
Total brain volume (25009)
Aseg derived total brain volume (26514)
Aseg derived brain volume not including the ventricles
Average cortical thickness for whole brain
Average of FA / MD / ICVF / ISOVF / OD / MO before scaling
```{r}

control_measures <- ukbio_IDP %>%
  select(n_eid, X25009.2.0, X26514.2.0, X26515.2.0, X26755.2.0, X26856.2.0)

```

Get the average cortical thickness over the whole brain by averaging left and right

IDPs 26755 and 26856 represent left and right "mean thickness of global mean thickness". We use them as a global variable alongside the aseg global variables
Included with the desikan IDPs, in desikan set

```{r}

control_measures <- control_measures %>%
  mutate(global_mean_thk = rowMeans(select(., X26755.2.0, X26856.2.0)))

```

Calculate columns of overall FA, MD, etc for each participant and add these to the control measures

```{r}

dMRI_means <- dMRI %>%
  pivot_longer(cols = -n_eid, names_to = "IDP", values_to = "value") %>%
  left_join(select(IDP_name_table, c(IDP, dMRI))) %>%
  select(-IDP) %>%
  group_by(n_eid, dMRI) %>%
  summarise(across(value, ~mean(.x, na.rm = TRUE))) %>%
  mutate(dMRI = paste0("mean_", dMRI)) %>%
  pivot_wider(id_cols = n_eid, names_from = dMRI, values_from = value)


```

```{r}

control_measures <- control_measures %>%
  left_join(dMRI_means, by = "n_eid")

```

Dont run na.omit on control measures, it has controls for multiple MRI types with different levels of missingness
The missingness in the control will match the IDPs they are used with


############################## Make a special subset for total volumes (will not be normalized)
Total volume IDPs
Units of measure for all of these is mm3
X25001
X25003
X25005
X25007
X25009
X25781 Total volume of white matter hyperintensities (from T1 and T2_FLAIR images) MISSING in set from Spring 2023
X26514 Volume of BrainSeg (whole brain)
26517 Volume of SubCortGray (whole brain)
X26528 Volume of WM-hypointensities (whole brain)
total_wm (from X26553.2.0 + X26584.2.0)
total_cortex (from X26552.2.0 + X26583.2.0)

Select IDPs to start the table
```{r}

total_volumes <- ukbio_IDP %>%
  select(c("n_eid", all_of(total_volume_IDPs[total_volume_IDPs %in% colnames(ukbio_IDP)]))) %>%
  mutate(global_mean_thk = rowMeans(select(., X26755.2.0, X26856.2.0))) %>%
  mutate(total_wm = X26553.2.0 + X26584.2.0) %>%
  mutate(total_cortex = X26552.2.0 + X26583.2.0) %>%
  left_join(dMRI_means, by = "n_eid") %>%
  # scale only the dMRI means
  mutate_at(vars(mean_FA, mean_MD, mean_ICVF, mean_ISOVF, mean_MO, mean_OD),
            ~scale(.x, center = TRUE, scale = TRUE))

```


Don't run na.omit on the total volumes, they have different levels of missingness. Use the n from the corresponding group of regional IDPs in the N tracker

############################## Scale IDPs

Scale all the IDP tables
```{r}

GMV <- mutate_at(GMV, vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))
free_2009 <- mutate_at(free_2009, vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))
desikan <- mutate_at(desikan, vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))
T1_regional <- mutate_at(T1_regional, vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))
T2 <- mutate_at(T2, vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))
dMRI <- mutate_at(dMRI, vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))
rfMRI <- mutate_at(rfMRI, vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))

# Also scale the control measures
control_measures <- mutate_at(control_measures, vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))

# Total volumes are NOT scaled (except dMRI measures), this way the output is in mm3 and more interpertable

```

After performing this transformation using mutate_at and scale, I checked the results for column X25011.2.0 by doing the calculation in excel. The column values match as of 9/20/21.
12/11/21 Also checked the results for the global columns after adding these.

#########Removal of outliers was here, final decision to drop this step after meeting with Janine on 5/20/22

Save the filtered and normalized IDPs tables
```{r}

save(GMV, file = paste0(output_path, "GMV_", Sys.Date(), ".RData"))
save(free_2009, file = paste0(output_path, "free_2009_", Sys.Date(), ".RData"))
save(desikan, file = paste0(output_path, "desikan_", Sys.Date(), ".RData"))
save(T1_regional, file = paste0(output_path, "T1_regional_", Sys.Date(), ".RData"))
save(T2, file = paste0(output_path, "T2_", Sys.Date(), ".RData"))
save(dMRI, file = paste0(output_path, "dMRI_", Sys.Date(), ".RData"))
save(rfMRI, file = paste0(output_path, "rfMRI_", Sys.Date(), ".RData"))

save(control_measures, file = paste0(output_path, "controls_", Sys.Date(), ".RData"))
write.csv(control_measures, file = paste0(output_path, "controls_", Sys.Date(), ".csv")) # As .csv becuse otherwise the columns are matrices

save(total_volumes, file = paste0(output_path, "total_volumes_", Sys.Date(), ".RData"))


```


Take the list of eid from ukbio IDP to select participants to include from imaging and non-imaging covariates
Why ukbio and not one of the subsets?
While some subsets (ex. GMV) have a lower rate of missing, all are missing some participants who are present in the others
5/6/22 bug fix: do remove every participant who is missing brain volume X25009.2.0
```{r}

filtered_eid <- ukbio_IDP %>%
  pull(n_eid)

save(filtered_eid, file = paste0(output_path, paste0("filtered_eid_p1_", Sys.Date(), ".RData")))

```


Save and view the n tracker

```{r}

save(n_tracker, file = paste0(output_path, "n_tracker_p1_", Sys.Date(), ".RData"))

n_tracker %>%
  pivot_wider(id_cols = MRI, names_from = Step, values_from = N) %>%
  mutate(across(everything(), ~formatC(.x, format = "d"))) %>%
  mutate(across(everything(), ~gsub("NA", "", .x))) %>%
  kable(caption = paste0("N Remaining by Processing Step")) %>%
  kable_classic(full_width = F, html_font = "Cambria")


```

5/6/22 Based on this table, it appears the bug fix is working and the 4 people who are missing brain volume are being dropped
6/28/23 No more people randomly missing total brain volume


```{r}

print(paste0("Script last run ", Sys.Date()))

```

