---
title: "Pipeline Step 5 - Covariates, backfilling, and imputation"
output:
  html_document:
    df_print: paged
---



```{r}
library(dplyr) # Run install.packages("tidyverse") to get this
library(tidyr) # Run install.packages("tidyr") to get this
library(ggplot2)
library("reshape") # Run install.packages("reshape") to get this
library("broom")
library(kableExtra) # Run install.packages("kableExtra")
library(mice)

```

Set the filepath to get input files
```{r}
# Home computer
input_path <- "file_path"

```

Set the filepath to save outputs to
```{r}
# Home computer
output_path <- "file_path"

```

Set the date the IDP names table script was run
```{r}

run_date <- "yyyy-mm-dd"

```


Load eid, missing report, alcohol, smoking, and imaging covariates (not the processed ones)
```{r}

load(paste0(output_path, "filtered_eid_p2_", run_date, ".RData"))
load(paste0(output_path, "missing_report_p4_", run_date, ".RData"))
load(paste0(output_path, "imaging_alcohol_", run_date, ".RData"))
load(paste0(output_path, "baseline_smoking_", run_date, ".RData"))
load(paste0(output_path, "imaging_smoking_", run_date, ".RData"))
load(paste0(output_path, "imaging_covariates_", run_date, ".RData"))

```

4/25/24 Note that "sex" in imaging covariates comes from Data field 22001 Genetic sex


1/2/24 Added the health status measures which are used to filter participants for train / test
These will be included in all regression models
2188 - Long-standing illness, disability or infirmity
2178 - Overall health rating
```{r}

health_status <- read.csv(paste0(input_path, "health_status.csv"))

```

Use this to create a column called healthy with "healthy" or "non-healthy"
Primary phenotype is the imaging visit, backfilled with baseline
Cole also references diabetes and stroke history

2178 uses data coding 100508
"In general how would you rate your overall health?"
1 = excellent, 2 = good, other values are less good or no answer

2188 uses data coding 100349
"Do you have any long-standing illness, disability or infirmity?"
1 = Yes, 0 = No

Diabetes (Data field 2443)
"Has a doctor ever told you that you have diabetes?"
Data coding 100349 (for data field 2443)
1	Yes
0	No
-1	Do not know
-3	Prefer not to answer

Age at first stroke (Data field 4056)
"What was your age when the stroke was first diagnosed?"
NA responses mean no diagnosis of stroke
-1 represents "Do not know"
-3 represents "Prefer not to answer"


```{r}

health_status <- health_status %>%
  filter(n_eid %in% filtered_eid) %>%
  select(n_eid, n_2178_2_0, n_2178_0_0,
                n_2188_2_0, n_2188_0_0,
                n_2443_2_0, n_2443_0_0,
                n_4056_2_0, n_4056_0_0) %>%
  # Replace invalid responses with NA, -1 and -3 are invalid
  mutate_at(vars(n_2178_2_0, n_2178_0_0,
                n_2188_2_0, n_2188_0_0,
                n_2443_2_0, n_2443_0_0), ~ifelse(.x < 0, NA, .x)) %>%
  # NA for stroke means no stroke diagnosis, -3 and -1 are invalid responses
  mutate(n_4056_2_0 = ifelse(is.na(n_4056_2_0), 0, n_4056_2_0)) %>%
  mutate(n_4056_0_0 = ifelse(is.na(n_4056_0_0), 0, n_4056_0_0)) %>%
  mutate(n_4056_2_0 = ifelse(n_4056_2_0 < 0, NA, n_4056_2_0)) %>%
  mutate(n_4056_0_0 = ifelse(n_4056_0_0 < 0, NA, n_4056_0_0)) %>%
  mutate(n_4056_2_0 = ifelse(n_4056_2_0 > 0, 1, n_4056_2_0)) %>%
  mutate(n_4056_0_0 = ifelse(n_4056_0_0 > 0, 1, n_4056_0_0)) %>%
  dplyr::rename("health_rating" = n_2178_2_0, "longstanding" = n_2188_2_0, "diabetes" = n_2443_2_0,
                "stroke" = n_4056_2_0)

# Tried to use case_when and it just breaks on stroke, not sure why
# mutate_at also breaks

```

4/25/24 Andrey suggested controlling for the presence of psychiatric disorders
Referenced data field 20002. Same procedure as for neurological disorders, but we do not filter them out. Just note that they have a psych condition.
Include the disease list from both baseline and imaging appointments, they are different
Participants filtered out if the disease is reported at any appointment

```{r}

disease_baseline <- read.csv(paste0(input_path, "disease_baseline.csv"))
disease_imaging <- read.csv(paste0(input_path, "disease_imaging.csv"))

```

These data fields include all disease codes, not just psychiatric

Combine baseline and imaging, filter for only the people in our imaging cohort
```{r}

all_disease <- disease_baseline %>%
  left_join(disease_imaging, by = "n_eid") %>%
  filter(n_eid %in% filtered_eid)

```

Codes for pyschiatric diseases
From Data Coding 6 https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=6

1286 = depression
1287 = anxiety / panic attacks
1288 = nervous breakdown
1289 = schizophrenia
1291 = mania / bipolar / manic depression

```{r}

psych_codes <- data.frame((rbind(
  c("Depression", 1286),
  c("Anxiety / panic attacks", 1287),
  c("Nervous breakdown", 1288),
  c("Schizophrenia", 1289),
  c("Mania / bipolar / manic depression", 1291))))

psych_codes <- psych_codes %>%
  dplyr::rename("condition" = X1, "code" = X2) %>%
  mutate(code = formatC(code, format = "d"))

```

Use the table of psych codes to construct a table for psychiatric disease
```{r}

psych_disease <- all_disease %>%
  pivot_longer(cols = -n_eid, names_to = "field", values_to = "code") %>%
  select(-field) %>%
  filter(! is.na(code)) %>%
  # Remove any duplicates due to data from both the baseline and imaging visits
  distinct(.keep_all = TRUE) %>%
  mutate(code = formatC(code, format = "d")) %>%
  filter(code %in% psych_codes$code) %>%
  left_join(psych_codes, by = "code")

```

Use the table to tally up how many total individuals are affected, and which diseases
```{r}

psych_disease_table <- psych_disease %>%
  group_by(condition) %>%
  tally() %>%
  left_join(psych_codes, by = "condition") %>%
  relocate(code, .before = n) %>%
  mutate(percent = n / length(filtered_eid) * 100) %>%
  mutate(percent = formatC(percent, digits = 2, format = "f")) %>%
  arrange(desc(n))

psych_disease_table %>%
  kable(caption = paste0("Prevalence of psychological conditions in Imaging cohort\nN=",
                         length(filtered_eid), ", N affected=", length(unique(psych_disease$n_eid)))) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Save the psych disease table, it will go in the supplemental info
```{r}

save(psych_disease_table, file = paste0(output_path, "psych_disease_table_", Sys.Date(), ".RData"))

```



Import the non-imaging covariates
Updated non-imaging covariates on 1/19/22
Updated and added waist / hip ratio on 3/14/22
Added stress (6145), physical activity (884 and 904), diabetes (2443), vascular / heart (6150), cancer (2453), other diagnoses (2473) on 6/27/23

1/2/24 Will use stroke from health status rather than vascular / heart
```{r}

non_imaging_covariates <- read.csv(paste0(input_path, "non_imaging_covariates.csv"))

```


Filter the non-imaging covariates by eid for those that will be included in regression
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  filter(non_imaging_covariates$n_eid %in% filtered_eid)

```

Rename the non-imaging covariates that will be included in regression so that they are human-readable
Also create the smoking_freq field
Do not rename diastolic BP here, it will get done when a new column is created
1/9/22 Removed the smoking freq column, re-did the smoking to match Yoonhoo's analysis more closely
1/20/22 Taking the imaging visit values, rather than the baseline visit
```{r}
non_imaging_covariates <- non_imaging_covariates %>%
  dplyr::rename("BMI" = n_21001_2_0) %>%
  dplyr::rename("income" = n_738_2_0) %>%
  dplyr::rename("weight" = n_21002_2_0) %>%
  dplyr::rename("brain_volume" = n_25009_2_0) %>%
  dplyr::rename("stress" = n_6145_2_0) %>%
  dplyr::rename("moderate" = n_884_2_0) %>%
  dplyr::rename("vigorous" = n_904_2_0) %>%
  dplyr::rename("cancer" = n_2453_2_0) %>%
  dplyr::rename("other_diagnosis" = n_2473_2_0)

```

Join with health status
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  select(-n_2443_0_0) %>% # prevent duplication of baseline diabetes
  left_join(health_status, by = "n_eid")

```

Make a column for psychiatric conditions using psych disease
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(psychiatric = ifelse(n_eid %in% psych_disease$n_eid, 1, 0))

non_imaging_covariates %>%
  group_by(psychiatric) %>%
  tally()

```


3/19/22 Added systolic BP

```{r}

missing_report <- rbind(missing_report, data.frame("step" = "Import", "data_field" = c("diastolic_BP", "systolic_BP"),
                             "n_missing" = c(sum(is.na(non_imaging_covariates$n_4079_2_0)),
                                             sum(is.na(non_imaging_covariates$n_4080_2_0)))))

```

BLOOD PRESSURE

Diastolic  BP
Data field 4079 is the machine blood pressure. When this could not be taken for any reason, a manual BP was obtained.
When the machine BP is missing, replace with the manual value
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(diastolic_BP = ifelse(is.na(n_4079_2_0), n_94_2_0, n_4079_2_0)) %>%
  mutate(baseline_DBP = ifelse(is.na(n_4079_0_0), n_94_0_0, n_4079_0_0))

```

Systolic BP
Data field 4080 = machine, 93 = manual
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(systolic_BP = ifelse(is.na(n_4080_2_0), n_93_2_0, n_4080_2_0)) %>%
  mutate(baseline_SBP = ifelse(is.na(n_4080_0_0), n_93_0_0, n_4080_0_0))

```


Update the number of BP measures missing for the imaging visit

```{r}

missing_report <- rbind(missing_report, data.frame("step" = "Manual BP", "data_field" = c("diastolic_BP", "systolic_BP"),
                             "n_missing" = c(sum(is.na(non_imaging_covariates$diastolic_BP)),
                             sum(is.na(non_imaging_covariates$systolic_BP)))))

```

WAIST HIP RATIO

Calculate waist hip ratio from the waist and hip measurements
Data field 48 = waist circumference
Data field 49 = hip circumference
Adding this was based on Gray 2020, they found average waist / hip as 0.8 which is consistent with dividing waist circumference by hip circumference

```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(waist_hip = n_48_2_0 / n_49_2_0) %>%
  mutate(baseline_WH = n_48_0_0 / n_49_0_0)

median(non_imaging_covariates$waist_hip, na.rm = TRUE)

```
The median I found is similar

No conversions needed for BMI, it is already continuous numeric

INCOME (Data field 738)
Data coding 100294
1	Less than 18,000
2	18,000 to 30,999
3	31,000 to 51,999
4	52,000 to 100,000
5	Greater than 100,000
-1	Do not know
-3	Prefer not to answer

Convert the income into ordered numeric values by taking the lower bound of each income range

```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(income = case_when(
    income == 1 ~ 0,
    income == 2 ~ 18000,
    income == 3 ~ 31000,
    income == 4 ~ 52000,
    income == 5 ~ 100000))
# Do not know, prefer not to answer, and missing all converted to NA

```

Repeat with the baseline
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(baseline_income = case_when(
    n_738_0_0 == 1 ~ 0,
    n_738_0_0 == 2 ~ 18000,
    n_738_0_0 == 3 ~ 31000,
    n_738_0_0 == 5 ~ 52000,
    n_738_0_0 == 5 ~ 100000))

```


EDUCATION

At one point we were using the "age completed education" variable but this was not collected for anyone with a college degree.
The data for education credentials is MUCH more complete.
How to convert these to numeric values for inclusion in regression? Years of education represented by the qualification is a good conversion.

For the non college folks, how well does the age completed education fill in the number of years they spent in school?
Per Zhou 2021, this can be done by comparing to International Standard Classification of Education
This conversion is spelled out in the supplementary info for the paper.

None of the above (no qualifications) = 7 years
CSEs or equivalent = 10 years
O levels/GCSEs or equivalent = 10 years
A levels/AS levels or equivalent = 13 years
Other professional qualification = 15 years
NVQ or HNC or equivalent = 19 years
College or university degree = 20 years

6/27/23 UKB added Data coding 100305
1	College or University degree
2	A levels/AS levels or equivalent
3	O levels/GCSEs or equivalent
4	CSEs or equivalent
5	NVQ or HND or HNC or equivalent
6	Other professional qualifications eg: nursing, teaching
-7	None of the above
-3	Prefer not to answer


For now I think this is sufficient and makes sense, although it is to be noted I know very little about UK education levels.
1/20/22 Updated to use the imaging visit value
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(education_years = case_when(
    n_6138_2_0 == -7 ~ 7,
    n_6138_2_0 == 4 ~ 10,
    n_6138_2_0 == 3 ~ 10,
    n_6138_2_0 == 2 ~ 13,
    n_6138_2_0 == 6 ~ 15,
    n_6138_2_0 == 5 ~ 19,
    n_6138_2_0 == 1 ~ 20))


```

Convert the baseline as well for backfilling

```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(baseline_education = case_when(
    n_6138_0_0 == -7 ~ 7,
    n_6138_0_0 == 4 ~ 10,
    n_6138_0_0 == 3 ~ 10,
    n_6138_0_0 == 2 ~ 13,
    n_6138_0_0 == 6 ~ 15,
    n_6138_0_0 == 5 ~ 19,
    n_6138_0_0 == 1 ~ 20))

```

STRESS (Data field 6145)
"In the last 2 years have you experienced any of the following? (You can select more than one answer)"

Data coding 100502
1	Serious illness, injury or assault to yourself
2	Serious illness, injury or assault of a close relative
3	Death of a close relative
4	Death of a spouse or partner
5	Marital separation/divorce
6	Financial difficulties
-7	None of the above
-3	Prefer not to answer

For now I am coding this as binary, people have had stress in the past two years or not

```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(stress = case_when(
    is.na(stress) ~ -3,
    stress == -3 ~ -3,
    stress == -7 ~ 0,
    TRUE ~ 1)) %>%
  mutate(stress = ifelse(stress == -3, NA, stress))

```

```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(baseline_stress = case_when(
    is.na(n_6145_0_0) ~ -3,
    n_6145_0_0 == -3 ~ -3,
    n_6145_0_0 == -7 ~ 0,
    TRUE ~ 1)) %>%
  mutate(n_6145_0_0 = ifelse(n_6145_0_0 == -3, NA, n_6145_0_0))

```

Moderate physical activity (Data field 884)

"In a typical WEEK, on how many days did you do 10 minutes or more of moderate physical activities like carrying light loads, cycling at normal pace? (Do not include walking)"
Values range from 0 to 7

Vigorous physical activity (Data field 904)

"In a typical WEEK, how many days did you do 10 minutes or more of vigorous physical activity? (These are activities that make you sweat or breathe hard such as fast cycling, aerobics, heavy lifting)"

Data coding 100291 for values outside of 0 to 7 for both 884 and 904
-1	Do not know
-3	Prefer not to answer
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  # Moderate
  mutate(moderate = ifelse(moderate < 0, NA, moderate)) %>%
  mutate(baseline_moderate = ifelse(n_884_0_0 < 0, NA, moderate)) %>%
  # Vigorous
  mutate(vigorous = ifelse(vigorous < 0, NA, vigorous)) %>%
  mutate(baseline_vigorous = ifelse(n_904_0_0 < 0, NA, vigorous))

```


PRIOR DIAGNOSES

cancer (Data field 2453)
"Has a doctor ever told you that you have had cancer?"

other diagnosis (Data field 2473)
"Has a doctor ever told you that you have had any other serious medical conditions or disabilities?"

Data coding 100603 (applied for both 2453 and 2473)
1	Yes - you will be asked about this later by an interviewer
0	No
-1	Do not know
-3	Prefer not to answer

For all these fields, <0 means an invalid response or none of the above, 0 is no, and > 0 means the patient has the relevant condition
For now I am coding these as binary variables, although this can be revisited to be more specific with the heart conditions
```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate_at(c("cancer", "other_diagnosis"), ~case_when(
    .x == -7 ~ 0,
    .x < 0 ~ -3,
    .x == 0 ~ 0,
    .x > 0 ~ 1)) %>%
  mutate_at(c("cancer", "other_diagnosis"), ~ifelse(.x == -3, NA, .x))

```


```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  dplyr::rename("baseline_cancer" = n_2453_0_0, "baseline_other" = n_2473_0_0) %>%
  mutate_at(c("baseline_cancer", "baseline_other"), ~case_when(
    .x == -7 ~ 0,
    .x < 0 ~ -3,
    .x == 0 ~ 0,
    .x > 0 ~ 1)) %>%
  mutate_at(c("baseline_cancer", "baseline_other"), ~ifelse(.x == -3, NA, .x))

```


Backfill all the non-imaging covariates (except for smoking, that is the next step)
Get counts of missing in all the imaging covariates

```{r}

missing_report <- rbind(missing_report, data.frame("step" = rep("Before", 18),
            "data_field" = c("diastolic_BP", "systolic_BP", "waist_hip", "BMI",
                             "income", "education_years", "brain_volume", "weight", "stress", "moderate_activity",
                             "vigorous_activity", "cancer", "other_diagnosis",
                             "health_rating", "longstanding", "diabetes", "stroke", "psychiatric"),
                            "n_missing" = c(sum(is.na(non_imaging_covariates$diastolic_BP)),
                                            sum(is.na(non_imaging_covariates$systolic_BP)),
                                            sum(is.na(non_imaging_covariates$waist_hip)),
                                            sum(is.na(non_imaging_covariates$BMI)),
                                            sum(is.na(non_imaging_covariates$income)),
                                            sum(is.na(non_imaging_covariates$education_years)),
                                            sum(is.na(non_imaging_covariates$brain_volume)),
                                            sum(is.na(non_imaging_covariates$weight)),
                                            sum(is.na(non_imaging_covariates$stress)),
                                            sum(is.na(non_imaging_covariates$moderate)),
                                            sum(is.na(non_imaging_covariates$vigorous)),
                                            sum(is.na(non_imaging_covariates$cancer)),
                                            sum(is.na(non_imaging_covariates$other_diagnosis)),
                                            sum(is.na(non_imaging_covariates$health_rating)),
                                            sum(is.na(non_imaging_covariates$longstanding)),
                                            sum(is.na(non_imaging_covariates$diabetes)),
                                            sum(is.na(non_imaging_covariates$stroke)),
                                            sum(is.na(non_imaging_covariates$psychiatric)))))

```

Now backfill with values from the baseline survey

```{r}

non_imaging_covariates <- non_imaging_covariates %>%
  mutate(diastolic_BP = ifelse(is.na(diastolic_BP), baseline_DBP, diastolic_BP)) %>%
  mutate(systolic_BP = ifelse(is.na(systolic_BP), baseline_SBP, systolic_BP)) %>%
  mutate(waist_hip = ifelse(is.na(waist_hip), baseline_WH, waist_hip)) %>%
  mutate(BMI = ifelse(is.na(BMI), n_21001_0_0, BMI)) %>%
  mutate(income = ifelse(is.na(income), baseline_income, income)) %>%
  mutate(education_years = ifelse(is.na(education_years), baseline_education, education_years)) %>%
  mutate(weight = ifelse(is.na(weight), n_21002_0_0, weight)) %>%
  mutate(stress = ifelse(is.na(stress), baseline_stress, stress)) %>%
  mutate(moderate = ifelse(is.na(moderate), baseline_moderate, moderate)) %>%
  mutate(vigorous = ifelse(is.na(vigorous), baseline_vigorous, vigorous)) %>%
  mutate(cancer = ifelse(is.na(cancer), baseline_cancer, cancer)) %>%
  mutate(other_diagnosis = ifelse(is.na(other_diagnosis), baseline_other, other_diagnosis)) %>%
  mutate(health_rating = ifelse(is.na(health_rating), n_2178_0_0, health_rating)) %>%
  mutate(longstanding = ifelse(is.na(longstanding), n_2188_0_0, longstanding)) %>%
  mutate(diabetes = ifelse(is.na(diabetes), n_2443_0_0, diabetes)) %>%
  mutate(stroke = ifelse(is.na(stroke), n_4056_0_0, stroke))
# No backfilling on psychiatric
  
# Not doing this for brain volume, as there is no value to backfill!
  
```


```{r}

missing_report <- rbind(missing_report, data.frame("step" = rep("After", 18),
            "data_field" = c("diastolic_BP", "systolic_BP", "waist_hip", "BMI",
                             "income", "education_years", "brain_volume", "weight", "stress", "moderate_activity",
                             "vigorous_activity", "cancer", "other_diagnosis",
                             "health_rating", "longstanding", "diabetes", "stroke", "psychiatric"),
                            "n_missing" = c(sum(is.na(non_imaging_covariates$diastolic_BP)),
                                            sum(is.na(non_imaging_covariates$systolic_BP)),
                                            sum(is.na(non_imaging_covariates$waist_hip)),
                                            sum(is.na(non_imaging_covariates$BMI)),
                                            sum(is.na(non_imaging_covariates$income)),
                                            sum(is.na(non_imaging_covariates$education_years)),
                                            sum(is.na(non_imaging_covariates$brain_volume)),
                                            sum(is.na(non_imaging_covariates$weight)),
                                            sum(is.na(non_imaging_covariates$stress)),
                                            sum(is.na(non_imaging_covariates$moderate)),
                                            sum(is.na(non_imaging_covariates$vigorous)),
                                            sum(is.na(non_imaging_covariates$cancer)),
                                            sum(is.na(non_imaging_covariates$other_diagnosis)),
                                            sum(is.na(non_imaging_covariates$health_rating)),
                                            sum(is.na(non_imaging_covariates$longstanding)),
                                            sum(is.na(non_imaging_covariates$diabetes)),
                                            sum(is.na(non_imaging_covariates$stroke)),
                                            sum(is.na(non_imaging_covariates$psychiatric)))))

```

Make a nice figure of the missing report to see the effect of backfilling

```{r}

missing_report %>%
  pivot_wider(id_cols = data_field, names_from = step, values_from = n_missing) %>%
  mutate(percent_baseline = (Before - After) / nrow(non_imaging_covariates) * 100) %>%
  mutate(across(-percent_baseline, ~formatC(.x, format = "d"))) %>%
  mutate(across(percent_baseline, ~formatC(.x, digits = 1, format = "f"))) %>%
  mutate(across(everything(), ~gsub("NA", "", .x))) %>%
  relocate(Import, .before = Before) %>%
  relocate(`Manual BP`, .before = Before) %>%
  kable(caption = paste0("N missing by analysis step (N=", nrow(non_imaging_covariates), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Systolic and diastolic BP have the same number missing, this makes sense since you get both at the same time
However, just to be 100% sure I haven't swapped them somewhere compare the medians
```{r}

print(paste0("Median systolic BP = ", median(non_imaging_covariates$systolic_BP, na.rm = TRUE)))
print(paste0("Median diastolic BP = ", median(non_imaging_covariates$diastolic_BP, na.rm = TRUE)))

```
These medians make sense

=========================================================================
Smoking

Updated 6/27/23 to work with the new data coding system in the UK Biobank data we got in Spring 2023. I reduced the number of variables to just ever_daily_smoker, never_smoker, and pack years since I have not been using the other smoking variables.

Updated on 1/9/22 to more closely match how Yoonhoo handles the flow of smoking questions.
Already renamed pack years. Initially many participants are missing pack years. Some are non-smokers, and there is also a group of light smokers who were not asked about onset / offset of smoking
This second group is MNAR, so cannot use imputation on them

As noted at the top of this script, this version assumes initial processing of the baseline and imaging smoking variables was done in additional_smoking_covariates.Rmd


Get the rate of missing in the imaging set before backfilling
```{r}

missing_report <- rbind(missing_report, data.frame("step" = rep("Before", 5),
                                                   "data_field" = c("ever_daily_smoked",
                                                                    "never_smoked",
                                                                    "smoking_status",
                                                                    "pack_years",
                                                                    "cig_per_day"),
                             "n_missing" = c(sum(is.na(imaging_smoking$ever_daily_smoked)),
                                             sum(is.na(imaging_smoking$never_smoked)),
                                             sum(is.na(imaging_smoking$smoking_status)),
                                             sum(is.na(imaging_smoking$pack_years)),
                                             sum(is.na(imaging_smoking$cig_per_day)))))

```


Combine the data sets

```{r}

# Rename the baseline columns so we can reference them
names(baseline_smoking) <- names(baseline_smoking) %>%
  paste0(., "_baseline")

# Fix eid
baseline_smoking <- baseline_smoking %>%
  dplyr::rename("n_eid" = n_eid_baseline)

all_smoking <- left_join(imaging_smoking, baseline_smoking, by = "n_eid")

```

Now backfill with values from the baseline survey

```{r}

all_smoking <- all_smoking %>%
  mutate(ever_daily_smoked = ifelse(is.na(ever_daily_smoked), ever_daily_smoked_baseline, ever_daily_smoked)) %>%
  mutate(never_smoked = ifelse(is.na(never_smoked), never_smoked_baseline, never_smoked)) %>%
  mutate(smoking_status = ifelse(is.na(smoking_status), smoking_status_baseline, smoking_status)) %>%
  mutate(pack_years = ifelse(is.na(pack_years), pack_years_baseline, pack_years)) %>%
  mutate(cig_per_day = ifelse(is.na(cig_per_day), cig_per_day_baseline, cig_per_day))

# Select relevant columns and remove the imaging / baseline data sets
all_smoking <- all_smoking %>%
  select(n_eid, ever_daily_smoked, never_smoked, smoking_status, pack_years, cig_per_day)

rm(imaging_smoking)
rm(baseline_smoking)


```

Update the missing report

```{r}

missing_report <- rbind(missing_report, data.frame("step" = rep("After", 5),
                                                   "data_field" = c("ever_daily_smoked",
                                                                    "never_smoked",
                                                                    "smoking_status",
                                                                    "pack_years",
                                                                    "cig_per_day"),
                             "n_missing" = c(sum(is.na(all_smoking$ever_daily_smoked)),
                                             sum(is.na(all_smoking$never_smoked)),
                                             sum(is.na(all_smoking$smoking_status)),
                                             sum(is.na(all_smoking$pack_years)),
                                             sum(is.na(all_smoking$cig_per_day)))))

```

Visualize the missing report again

```{r}

missing_report %>%
  pivot_wider(id_cols = data_field, names_from = step, values_from = n_missing) %>%
  relocate("Manual BP", .before = Before) %>%
  relocate(Import, .before = "Manual BP") %>%
  mutate(percent_backfilled = (Before - After) / nrow(non_imaging_covariates) * 100) %>%
  mutate(across(-percent_backfilled, ~formatC(.x, format = "d"))) %>%
  mutate(across(percent_backfilled, ~formatC(.x, digits = 1, format = "f"))) %>%
  mutate(across(everything(), ~gsub("NA", "", .x))) %>%
  kable(caption = paste0("N missing by analysis step (N=", nrow(non_imaging_covariates), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```
With the backfilling procedure there are very few missing values left!

Start the imputation table with the relevant columns from non-imaging covariates + alcohol + smoking

```{r}

imputation_table <- non_imaging_covariates %>%
  select(n_eid, diastolic_BP, systolic_BP, waist_hip, BMI, weight, income, education_years, brain_volume,
         stress, moderate, vigorous, cancer, other_diagnosis, health_rating, longstanding, diabetes, stroke, psychiatric) %>%
  left_join(imaging_alcohol, by = "n_eid")  %>%
  left_join(all_smoking, by = "n_eid") %>%
  left_join(select(imaging_covariates, c(n_eid, sex, date, age)), by = "n_eid")

```

Make a table with mean and SD of non-imaging covariates before imputation
Note that this is only done for the numeric variables

Set a list of the non-numeric variables
```{r}

non_numeric <- c("n_eid", "sex", "stress", "cancer", "other_diagnosis", "never_alcohol", "alcohol_status",
            "smoking_status", "longstanding", "diabetes", "stroke", "psychiatric", "ever_daily_smoked", "never_smoked")

```


```{r}

imputation_table %>%
  select(-all_of(non_numeric)) %>%
  summarise_all(funs(mean, sd, min, max), na.rm = TRUE) %>%
  pivot_longer(cols = everything(), names_to = "name", values_to = "value") %>%
  mutate(measure = case_when(
    grepl("_mean", name) ~ "mean",
    grepl("_sd", name) ~ "sd",
    grepl("_min", name) ~ "min",
    grepl("_max", name) ~ "max")) %>%
  mutate(name = gsub("_mean", "", name)) %>%
  mutate(name = gsub("_sd", "", name)) %>%
  mutate(name = gsub("_min", "", name)) %>%
  mutate(name = gsub("_max", "", name)) %>%
  pivot_wider(id_cols = name, names_from = measure, values_from = value) %>%
  mutate(mean = formatC(mean, digits = 2, format = "f")) %>%
  mutate(sd = formatC(sd, digits = 2, format = "f"))  %>%
  mutate(min = formatC(min, digits = 2, format = "f"))  %>%
  mutate(max = formatC(max, digits = 2, format = "f"))  %>%
  kable(caption = paste0("Mean and SD in Non-Imaging Covariates Before Imputation (N=",
                         nrow(non_imaging_covariates), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Save the imputation table (pre-scaling)
```{r}

save(imputation_table, file = paste0(output_path, "pre_imputation_covariates_", Sys.Date(), ".RData"))

```

Check the numbers of missing values by variable in the imputation table before performing imputation
```{r}

imputation_table %>%
  summarise_at(vars(-n_eid), ~ sum(is.na(.x))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "n_missing") %>%
  mutate(percent_missing = n_missing / nrow(imaging_covariates) * 100) %>%
  mutate(percent_missing = formatC(percent_missing, digits = 3, format = "f")) %>%
  kable(caption = paste0("Missing Values Before Imputation (N=", nrow(imputation_table), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

==================================Remove missing brain volume================================================

Because I started included a larger set up participants in the filtered eid, there are a few who are missing total brain volume. This cannot be backfilled, and in my opinion is too critical to the analysis to impute. So these participants should be dropped.

5/6/22 Bug fix should mean 0 participants are dropped here, but double check this each time.
```{r}

imputation_table <- imputation_table %>%
  filter(! is.na(brain_volume))

print(paste0("Dropped ", length(filtered_eid) - nrow(imputation_table), " participants for missing brain volume"))

```

=========================Normalize and handle skew============================================

The IDPs are z-score normalized, the imaging covariates are z-score normalized, so the alcohol / smoking phenotypes
and non-imaging covariates should be normalized as well.

Before normalization, capture the mean and sd of all the columns in order to de-normalize some measures later
Exclude columns which are binary (sex, stress, diabetes, cancer, other_diagnosis, longstanding, stroke,
ever_daily_smoked, never_smoked)
```{r}

means <- imputation_table %>%
  summarize(across(-all_of(non_numeric),
                   ~mean(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "covariate", values_to = "mean")

sd <- imputation_table %>%
  summarize(across(-all_of(non_numeric),
                   ~sd(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "covariate", values_to = "sd")

pre_scaling <- left_join(means, sd, by = "covariate")

save(pre_scaling, file = paste0(output_path, "pre_scaling_", Sys.Date(), ".RData"))

```


Normalize all the continuous / numeric measures
```{r}

imputation_table <- imputation_table %>%
  mutate(across(-all_of(non_numeric),
                ~scale(.x, center = TRUE, scale = TRUE)))

```


```{r}

imputation_table %>%
  select(!all_of(non_numeric)) %>%
  pivot_longer(cols = everything(), names_to = "covariate", values_to = "values")  %>%
  ggplot(aes(x = values)) +
  geom_histogram() +
  facet_wrap(~ covariate)


```

Removed log transforming drinks and pack years, they are skewed and that is ok

Save the table as a .csv
Why to do this? There is an error about some of the columns being matrices that cannot be fixed any other way.
R why are you like this?

```{r}

write.csv(imputation_table, file = paste0(output_path, "pre_imputation_", Sys.Date(), ".csv"))
rm(imputation_table)

```

==================================================================================================
IMPUTATION

Keep track of what type of variable the columns being imputed are:
  diastolic_BP = numeric
  systolic BP = numeric
  waist-hip = numeric
  BMI = numeric
  weight = numeric
  income = numeric
  education_years" = numeric
***stress = logical
  moderate = numeric
  vigorous = numeric
***diabetes = logical
***cancer = logical
***other_diagnosis = logical
***stoke = logical
***psychiatric = logical
  drink_freq = numeric
  week_drinks = numeric
***never_alcohol = logical
***alcohol_status = categorical
***ever_daily_smoked = logical
***never_daily_smoked = logical
***smoking_status = categorical
  pack_years = numeric
  cig_per_day = numeric
***sex = logical
  date = numeric
  age = numeric

Reload the imputation table - saving and reloading makes the columns NOT matrices and fixes the cannot handle matrix error
```{r}

imputation_table<- read.csv(paste0(output_path, "pre_imputation_", Sys.Date(), ".csv"))

imputation_table <- imputation_table %>%
  select(-X)

```


At one point I had written my own function for this
Now I am using the mice function so it will be more standard
```{r}

set.seed(1987)
post_imputation <- mice(imputation_table, m = 15, method = "cart", seed = 6)

```

Some of the variables do not show up in imputation, this is because they have no missing

Take a look at convergence in some of the categories with more missing values, is 5 enough rounds?

```{r}

post_imputation$imp$education_years %>%
  tibble::rownames_to_column(var = "n_eid") %>%
  mutate("1" = (1 - 2) / 1) %>%
  mutate("2" = (2 - 3) / 2) %>%
  mutate("3" = (3 - 4) / 3) %>%
  mutate("4" = (4 - 5) / 4) %>%
  mutate("5" = (5 - 6) / 5) %>%
  mutate("6" = (6 - 7) / 6) %>%
  mutate("7" = (7 - 8) / 7) %>%
  mutate("8" = (8 - 9) / 8) %>%
  mutate("9" = (9 - 10) / 9) %>%
  mutate("10" = (10 - 11) / 10) %>%
  mutate("11" = (11 - 12) / 11) %>%
  mutate("12" = (12 - 13) / 12) %>%
  mutate("13" = (13 - 14) / 13) %>%
  mutate("14" = (14 - 15) / 14) %>%
  select(n_eid, "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14") %>%
  pivot_longer(cols = -n_eid, names_to = "round", values_to = "value") %>%
  mutate(round = factor(round, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14"))) %>% {
  ggplot(., aes(x = round, y = value, color = n_eid, group = n_eid)) +
  geom_line() +
  geom_point(size = 3) +
  ggtitle(paste0("Convergence in Education Years (Missing=", length(unique(.$n_eid)), ")")) +
      ylim(c(-2, 0))
  }


```
Take a look at convergence in income, which has the most missing valeus

```{r}

post_imputation$imp$income %>%
  tibble::rownames_to_column(var = "n_eid") %>%
  mutate("1" = (1 - 2) / 1) %>%
  mutate("2" = (2 - 3) / 2) %>%
  mutate("3" = (3 - 4) / 3) %>%
  mutate("4" = (4 - 5) / 4) %>%
  mutate("5" = (5 - 6) / 5) %>%
  mutate("6" = (6 - 7) / 6) %>%
  mutate("7" = (7 - 8) / 7) %>%
  mutate("8" = (8 - 9) / 8) %>%
  mutate("9" = (9 - 10) / 9) %>%
  mutate("10" = (10 - 11) / 10) %>%
  mutate("11" = (11 - 12) / 11) %>%
  mutate("12" = (12 - 13) / 12) %>%
  mutate("13" = (13 - 14) / 13) %>%
  mutate("14" = (14 - 15) / 14) %>%
  select(n_eid, "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14") %>%
  pivot_longer(cols = -n_eid, names_to = "round", values_to = "value") %>%
  mutate(round = factor(round, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14"))) %>% {
  ggplot(., aes(x = round, y = value, color = n_eid, group = n_eid)) +
  geom_line() +
  geom_point(size = 3) +
  ggtitle(paste0("Convergence in Income (Missing=", length(unique(.$n_eid)), ")")) +
      ylim(c(-2, 0)) +
      theme(legend.position = "none")
  }



```

Save the full post imputation results
```{r}

save(post_imputation, file = paste0(output_path, "imputation_results_", Sys.Date(), ".RData"))

```
The above chunk might seem redundant but it's really helpful if you want to change small things in the file without re-running the whole imputation.

```{r}

imputation_results <- complete(post_imputation, 15)

```


Check to see that all missing have been filled in with an imputed value
```{r}

imputation_results %>%
  summarise_at(vars(-n_eid), ~ sum(is.na(.x))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "n_missing") %>%
  mutate(percent_missing = n_missing / nrow(imaging_covariates) * 100) %>%
  mutate(percent_missing = formatC(percent_missing, digits = 3, format = "f")) %>%
  arrange(desc(n_missing)) %>%
  kable(caption = paste0("Missing Values After Imputation (N=", nrow(imputation_table), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Alcohol status and smoking status do not always get imputed
Fix any that are still missing using other fields
As of 2/13/24, only 1 of the missing smoking status is not a never_smoked, so easy to assign most of them correctly
Former is the most likely smoking status for the remaining person
For alcohol status, if drink_freq > 0, then they are a current drinker
```{r}

drink_freq_scaled <- (0 - pull(select(filter(pre_scaling, covariate == "drink_freq"), mean)) / pull(select(filter(pre_scaling, covariate == "drink_freq"), sd)))

imputation_results <- imputation_results %>%
  # never_smoked == TRUE means smoking status should be Never
  mutate(smoking_status = ifelse(never_smoked == 1, "Never", smoking_status)) %>%
  # Former is most likely status if never_smoked == FALSE
  mutate(smoking_status = ifelse(is.na(smoking_status), "Former", smoking_status)) %>%
  # If drink_freq > 0, then alcohol status is Current
  mutate(alcohol_status = ifelse(drink_freq > drink_freq_scaled, "Current", alcohol_status))

imputation_results %>%
  summarise_at(vars(-n_eid), ~ sum(is.na(.x))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "n_missing") %>%
  mutate(percent_missing = n_missing / nrow(imaging_covariates) * 100) %>%
  mutate(percent_missing = formatC(percent_missing, digits = 3, format = "f")) %>%
  arrange(desc(n_missing))

```


=======================================================================================================
Post imputation

Make a version without scaling using the means and sd saved before scaling

```{r}

un_scaled <- imputation_results %>%
  select(-all_of(non_numeric[! non_numeric == "n_eid"])) %>%
  pivot_longer(cols = -n_eid, names_to = "covariate", values_to = "value") %>%
  left_join(pre_scaling, by = "covariate") %>%
  mutate(value = value * sd + mean) %>%
  select(-c(mean, sd)) %>%
  pivot_wider(id_cols = n_eid, names_from = covariate, values_from = value)

```

As a validation, take a look at the distributions, means, and sd after de-scaling to make sure they are broadly similar to the pre-imputation values

```{r}

un_scaled %>%
  select(-n_eid) %>%
  summarise_all(funs(mean, sd, min, max), na.rm = TRUE) %>%
  pivot_longer(cols = everything(), names_to = "name", values_to = "value") %>%
  mutate(measure = case_when(
    grepl("_mean", name) ~ "mean",
    grepl("_sd", name) ~ "sd",
    grepl("_min", name) ~ "min",
    grepl("_max", name) ~ "max")) %>%
  mutate(name = gsub("_mean", "", name)) %>%
  mutate(name = gsub("_sd", "", name)) %>%
  mutate(name = gsub("_min", "", name)) %>%
  mutate(name = gsub("_max", "", name)) %>%
  pivot_wider(id_cols = name, names_from = measure, values_from = value) %>%
  mutate(mean = formatC(mean, digits = 2, format = "f")) %>%
  mutate(sd = formatC(sd, digits = 2, format = "f"))  %>%
  mutate(min = formatC(min, digits = 2, format = "f"))  %>%
  mutate(max = formatC(max, digits = 2, format = "f"))  %>%
  kable(caption = paste0("Mean and SD in Non-Imaging Covariates After Imputation (N=",
                         nrow(non_imaging_covariates), ")")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Means and sd are largely the same comparing to the table from before imputation

Add the non-numeric variables to unscaled
```{r}

un_scaled <- un_scaled %>%
  left_join(select(imputation_results, all_of(non_numeric)), by = "n_eid")

```

Add the un-scaled columns for alcohol, pack years, and age to the main imputation results table
This makes running the regressions easy

```{r}

imputation_results <- imputation_results %>%
  dplyr::rename("scaled_week_drinks" = week_drinks) %>%
  dplyr::rename("scaled_pack_years" = pack_years) %>%
  dplyr::rename("scaled_cig_per_day" = cig_per_day) %>%
  dplyr::rename("scaled_age" = age) %>%
  left_join(select(un_scaled, c(n_eid, week_drinks, pack_years, cig_per_day, age)), by = "n_eid")

```


Make some additional alcohol columns

```{r}

# Divide drinks per month (month_drinks) by drinking days per month (drink_freq)
imputation_results <- imputation_results %>%
  mutate(drink_freq = drink_freq * pull(select(filter(pre_scaling, covariate == "drink_freq"), sd)) + 
           pull(select(filter(pre_scaling, covariate == "drink_freq"), mean))) %>%
  mutate(drinks_per_day = ifelse(drink_freq == 0, 0, (week_drinks * 4.3) / drink_freq))


# Create a binary_binge column - are typical drinks per drink day 5 or more?
imputation_results <- imputation_results %>%
  mutate(binary_binge = ifelse(drinks_per_day >= 5, 1, 0))

```


Save the final imputed covariates! Next up is the brain age model itself

```{r}

save(imputation_results, file = paste0(output_path, "imputed_covariates_", Sys.Date(), ".RData"))
save(un_scaled, file = paste0(output_path, "un_scaled_covariates_", Sys.Date(), ".RData"))
save(missing_report, file = paste0(output_path, "missing_report_p5_", Sys.Date(), ".RData"))

```


```{r}

print(paste0("Script last run ", Sys.Date()))

```




